---
title: "Elo-Rating"
output: html_document
date: "2024-08-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/lukeconkleton/Library/Mobile Documents/com~apple~CloudDocs/Documents/UCL Project/Data")
```

```{r, include=FALSE}
library(tidyverse)
library(brms)
library(coda)
library(reshape2)
```

#############################################################################
#############################################################################

```{r}
rounds_data <- readRDS("roundData2018-2022.rds")
odds_data <- readRDS("oddsData2018-2022.rds")
```

## 1. DATA WRANGLING & FEATURE SELECTION

Subsetting dataframe to only include pre-covid data. Creating a new dataframe to only inlcude variables of interest; golfer, tournament, major, final position and date.

```{r}
rounds_data <- rounds_data %>% filter(year %in% c('2018', '2019'))
```

Engineering a feature for whether or not the golfer is playing at 'home':

```{r}
rounds_data$is.home <- ifelse(rounds_data$player_country == rounds_data$tournament_country, 1, 0)
```

```{r}
rounds <- data.frame(
  EP_game_id = rounds_data$EP_game_id,
  tournament = rounds_data$event_name,
  major = rounds_data$is_major,
  golfer_id = rounds_data$player_id,
  golfer = rounds_data$player_name,
  final_position = rounds_data$final_position,
  start_date = rounds_data$competition_start_date, 
  times_played_course = rounds_data$times_played_course,
  is.home = rounds_data$is.home
)

rounds <- rounds %>%
  distinct()
```

Only keeping a variable for times played course pre-tournament.

```{r}
rounds <- rounds %>%
  group_by(EP_game_id, tournament, golfer_id, start_date) %>%
  slice_min(times_played_course) %>% #takes the minimum value of the datapoints
  ungroup() %>%
  rename(times_played_course_pre_tournament = times_played_course)
```



Only considering the relevant time frame:

```{r}
odds_data$competition_start_date <- as.Date(odds_data$competition_start_date)

odds_data <- odds_data %>% filter(format(competition_start_date, "%Y") %in% c("2018", "2019"))

odds_data <- odds_data %>%
  select(EP_game_id, event_name, SMARTODDS_golfer1_name, SMARTODDS_golfer2_name, golfer1_win, golfer2_win) %>%
  rename(
    player1 = SMARTODDS_golfer1_name,
    player2 = SMARTODDS_golfer2_name,
    tournament = event_name
  )
```

More tournaments included in rounds data set comparatively:

```{r}
length(unique(odds_data$EP_game_id))
length(unique(rounds_data$EP_game_id))
```

Subsetting the rounds dataset to only include golfers who have made over 50 appearances in the odds dataset:

```{r}
odds_golfers <- c(odds_data$player1, odds_data$player2)
length(unique(odds_golfers))

# how many appearances in the odds data frame
golfer_counts <- data.frame(golfer = odds_golfers) %>%
  group_by(golfer) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

players_over_50 <- golfer_counts %>%
  filter(count >= 50)

dim(players_over_50)
```


```{r}
# filter the dataframes to keep only those players
filtered_rounds_data <- rounds_data %>%
  filter(player_name %in% players_over_50$golfer)

filtered_rounds <- rounds %>%
  filter(golfer %in% players_over_50$golfer) %>%
  filter(!is.na(final_position) & final_position != 0)
```

## 2. CREATING A PAIRWISE COMPARISON DATASET

Create a dataset with every single match-up in that time period.


```{r}
# define the create_pairs function
create_pairs <- function(df) {
  pairs <- expand.grid(golfer_A = df$golfer_id, golfer_B = df$golfer_id)
  pairs <- pairs[pairs$golfer_A != pairs$golfer_B, ]
  
  golfer_A_indices <- match(pairs$golfer_A, df$golfer_id)
  golfer_B_indices <- match(pairs$golfer_B, df$golfer_id)
  
  if (length(golfer_A_indices) == 0 || length(golfer_B_indices) == 0) {
    return(data.frame()) 
  }
  
  result <- data.frame(
    golfer_A = df$golfer[golfer_A_indices],
    golfer_B = df$golfer[golfer_B_indices],
    final_position_A = df$final_position[golfer_A_indices],
    final_position_B = df$final_position[golfer_B_indices],
    start_date = df$start_date[golfer_A_indices],
    major = df$major[golfer_A_indices],
    EP_game_id = df$EP_game_id[golfer_A_indices],
    tournament = df$tournament[golfer_A_indices],
    playerA_times_played = df$times_played_course_pre_tournament[golfer_A_indices],
    playerB_times_played = df$times_played_course_pre_tournament[golfer_B_indices],
    is.home.A = df$is.home[golfer_A_indices],
    is.home.B = df$is.home[golfer_B_indices]
  )
  
  return(result)
}

# apply the function to each tournament
pairwise_matches <- filtered_rounds %>%
  group_by(EP_game_id) %>%
  do({
    result <- create_pairs(.)
    if (nrow(result) > 0) {
      result$EP_game_id <- .$EP_game_id[1]
    }
    result
  }) %>%
  ungroup()

pairwise_matches <- pairwise_matches %>%
  mutate(
    result_A = ifelse(final_position_A < final_position_B, 1, 0),
    result_B = ifelse(final_position_B < final_position_A, 1, 0)
  )

```

Ensure that there are no duplicate results in the dataset and that the data is ordered temporally.

```{r}
# ensure each pair is in a consistent order
pairwise_matches <- pairwise_matches %>%
  rowwise() %>%
  mutate(
    golfer_A_sorted = ifelse(golfer_A < golfer_B, golfer_A, golfer_B),
    golfer_B_sorted = ifelse(golfer_A < golfer_B, golfer_B, golfer_A),
    final_position_A_sorted = ifelse(golfer_A < golfer_B, final_position_A, final_position_B),
    final_position_B_sorted = ifelse(golfer_A < golfer_B, final_position_B, final_position_A),
    result_A_sorted = ifelse(golfer_A < golfer_B, result_A, result_B),
    result_B_sorted = ifelse(golfer_A < golfer_B, result_B, result_A),
    is.home_A_sorted = ifelse(golfer_A < golfer_B, is.home.A, is.home.B),
    is.home_B_sorted = ifelse(golfer_A < golfer_B, is.home.B, is.home.A),
    times_played_A_sorted = ifelse(golfer_A < golfer_B, playerA_times_played, playerB_times_played),
    times_played_B_sorted = ifelse(golfer_A < golfer_B, playerB_times_played, playerA_times_played)
  ) %>%
  ungroup()

# add home advantage columns after sorting
pairwise_matches <- pairwise_matches %>%
  mutate(
    home_advantage_A = ifelse(is.home_A_sorted == 1 & is.home_B_sorted == 0, 1, 0),
    home_advantage_B = ifelse(is.home_A_sorted == 0 & is.home_B_sorted == 1, 1, 0)
  )

# remove duplicates based on the sorted pairs and the tournament
pairwise_matches <- pairwise_matches %>%
  select(EP_game_id, tournament, golfer_A_sorted, golfer_B_sorted, final_position_A_sorted, final_position_B_sorted, 
         result_A_sorted, result_B_sorted, start_date, major, times_played_A_sorted, times_played_B_sorted, 
         is.home_A_sorted, is.home_B_sorted, home_advantage_A, home_advantage_B) %>%
  distinct() %>%
  rename(
    golfer_A = golfer_A_sorted,
    golfer_B = golfer_B_sorted,
    final_position_A = final_position_A_sorted,
    final_position_B = final_position_B_sorted,
    result_A = result_A_sorted,
    result_B = result_B_sorted,
    is.home_A = is.home_A_sorted,
    is.home_B = is.home_B_sorted,
    times_played_A = times_played_A_sorted,
    times_played_B = times_played_B_sorted
  )

pairwise_matches <- pairwise_matches %>%
  rename(
    player1 = golfer_A,
    player2 = golfer_B,
    score1 = result_A,
    score2 = result_B
  )

# maintaining temporal ordering
pairwise_matches <- pairwise_matches %>%
  arrange(start_date, EP_game_id)
```

#############################################################################
#############################################################################

# creating a predictive dataset

```{r}
tournament_summary <- pairwise_matches %>%
  group_by(tournament, EP_game_id, major, start_date) %>%
  summarise() %>%
  ungroup() %>%
  arrange(start_date, EP_game_id) %>%
  mutate(
    next_EP_game_id = lead(EP_game_id),
    next_tournament = lead(tournament),
    next_major = lead(major),
    next_start_date = lead(start_date),
    prev_EP_game_id = lag(EP_game_id),
    prev_tournament = lag(tournament),
    prev_major = lag(major),
    prev_start_date = lag(start_date)
  ) %>%
  filter(!is.na(next_EP_game_id) | row_number() == n()) 

dim(tournament_summary)
```

## CREATE A TRAINING SET

```{r}
unique_tournaments <- unique(pairwise_matches$EP_game_id)
length(unique_tournaments)
```

```{r}
training_set_size <- 80

create_training_set <- function(data, tournaments, num_tournaments) {
  training_tournaments <- tournaments[1:num_tournaments]
  training_data <- data[data$EP_game_id %in% training_tournaments, ]
  return(training_data)
}

training_set <- create_training_set(pairwise_matches, unique_tournaments, training_set_size)
```


## MODEL SET UP

Update schedule:

```{r}
# Initialize ratings for each player using the training set
players <- unique(c(training_set$player1, training_set$player2))
initial_rating <- 1500
elo_ratings <- data.frame(player = players, rating = initial_rating, stringsAsFactors = FALSE)
```

```{r}
# Function to update Elo ratings
update_elo <- function(ratingA, ratingB, scoreA, k = 32) {
    expectedA <- 1 / (1 + 10^((ratingB - ratingA) / 400))
    delta <- k * (scoreA - expectedA)
    return(delta)
}

# Calculate Elo ratings for the training set
for (i in 1:nrow(training_set)) {
    player1 <- training_set$player1[i]
    player2 <- training_set$player2[i]
    score1 <- training_set$score1[i]
    score2 <- training_set$score2[i]

    # Get current ratings
    rating1 <- elo_ratings$rating[elo_ratings$player == player1]
    rating2 <- elo_ratings$rating[elo_ratings$player == player2]

    # Calculate the rating change (delta)
    delta <- update_elo(rating1, rating2, score1)

    # Update ratings symmetrically
    new_rating1 <- rating1 + delta
    new_rating2 <- rating2 - delta

    # Assign new ratings
    elo_ratings$rating[elo_ratings$player == player1] <- new_rating1
    elo_ratings$rating[elo_ratings$player == player2] <- new_rating2
}
```

Initialising a data frame to store Elo-ratings and sets up functions to retrieve data following each tournament:

```{r}
initial_elo_ratings <- elo_ratings
strengths_wide <- data.frame(Player = players, stringsAsFactors = FALSE)

# get the last game ID from the training set
last_game_id <- last(training_set$EP_game_id)

# add the initial column with heading as the last game ID and containing initial strengths
strengths_wide[[as.character(last_game_id)]] <- initial_elo_ratings$rating[match(players, initial_elo_ratings$player)]

# add columns for each subsequent tournament
unique_tournaments <- unique(pairwise_matches$EP_game_id)
for (tournament_id in unique_tournaments[(training_set_size + 1):length(unique_tournaments)]) {
  strengths_wide[[as.character(tournament_id)]] <- NA
}

# initialize the current skills with the final training set ratings
current_skills <- initial_elo_ratings

get_new_tournament_data <- function(data, tournament_id) {
  data[data$EP_game_id == tournament_id, ]
}

```

Tracking and updating Elo ratings over tournaments, storing ratings in a wide format.

```{r}
# loop through tournaments after the initial training set
for (tournament_id in unique_tournaments[(training_set_size + 1):length(unique_tournaments)]) {
    new_tournament_data <- get_new_tournament_data(pairwise_matches, tournament_id)
    
    # update skills for each match in the tournament
    for (j in 1:nrow(new_tournament_data)) {
        player1 <- new_tournament_data$player1[j]
        player2 <- new_tournament_data$player2[j]
        score1 <- new_tournament_data$score1[j]
        score2 <- new_tournament_data$score2[j]

        # get current ratings, initializing if not present using initial_elo_ratings
        if (!player1 %in% current_skills$player) {
            initial_rating1 <- initial_elo_ratings$rating[initial_elo_ratings$player == player1]
            current_skills <- rbind(current_skills, data.frame(player = player1, rating = initial_rating1))
        } else {
            initial_rating1 <- current_skills$rating[current_skills$player == player1]
        }

        if (!player2 %in% current_skills$player) {
            initial_rating2 <- initial_elo_ratings$rating[initial_elo_ratings$player == player2]
            current_skills <- rbind(current_skills, data.frame(player = player2, rating = initial_rating2))
        } else {
            initial_rating2 <- current_skills$rating[current_skills$player == player2]
        }

        # update ratings symmetrically
        delta <- update_elo(initial_rating1, initial_rating2, score1)
        new_rating1 <- initial_rating1 + delta
        new_rating2 <- initial_rating2 - delta

        # assign new ratings
        current_skills$rating[current_skills$player == player1] <- new_rating1
        current_skills$rating[current_skills$player == player2] <- new_rating2
    }

    # store current skills after each update in wide format
    for (player in players) {
        if (player %in% current_skills$player) {
            current_strength <- current_skills$rating[current_skills$player == player]
            strengths_wide[strengths_wide$Player == player, as.character(tournament_id)] <- current_strength
        } else {
            previous_tournament_id <- as.character(unique_tournaments[which(unique_tournaments == tournament_id) - 1])
            strengths_wide[strengths_wide$Player == player, as.character(tournament_id)] <- strengths_wide[strengths_wide$Player == player, previous_tournament_id]
        }
    }
}
```

Check the zero-sum property of strengths dataframe:

```{r}
48*1500 # each column should equal this

sum(strengths_wide$"2915586")
sum(strengths_wide$"3096661")
```



## PREDICTION ANALYSIS

Taking out irrelevant columns:

```{r}
pairwise_matches <- pairwise_matches %>%
  select(-final_position_A, -final_position_B, -score2, -times_played_A, -times_played_B, -is.home_A, -is.home_B) %>%
  rename(outcome = score1)
```

Creating predictive dataframe:

```{r}
predictive_dataframe_e <- pairwise_matches %>%
  left_join(tournament_summary %>% select(EP_game_id, prev_EP_game_id, prev_tournament, prev_major, prev_start_date), by = "EP_game_id")
```


Merging Elo-ratings:

```{r}
strengths_long <- strengths_wide %>%
  pivot_longer(cols = -Player, names_to = "EP_game_id", values_to = "strength") %>%
  mutate(EP_game_id = as.numeric(EP_game_id))

# first join for player1 strengths
predictive_dataframe_e <- predictive_dataframe_e %>%
  left_join(strengths_long, by = c("prev_EP_game_id" = "EP_game_id", "player1" = "Player")) %>%
  rename(strength_player1 = strength)

names(predictive_dataframe_e)[names(predictive_dataframe_e) == "strength_player1"] <- "strength_player1_temp"

# second join for player2 strengths
predictive_dataframe_e <- predictive_dataframe_e %>%
  left_join(strengths_long, by = c("prev_EP_game_id" = "EP_game_id", "player2" = "Player")) %>%
  rename(strength_player2 = strength)

# restore the original name for strength_player1
names(predictive_dataframe_e)[names(predictive_dataframe_e) == "strength_player1_temp"] <- "strength_player1"
```

Adding columns model probabilities:

```{r}
predictive_dataframe_e <- predictive_dataframe_e %>%
  mutate(prob_player1_wins = 1 / (1 + 10^((strength_player2 - strength_player1) / 400)))

predictive_dataframe_e <- predictive_dataframe_e %>%
  mutate(prob_player2_wins = 1 - prob_player1_wins)
```

Getting rid of all non-predicitve data:

```{r}
predictive_dataframe_e <- predictive_dataframe_e %>%
  drop_na()
```

Accuracy of this dataset? Pointless

```{r}
correct_predictions <- predictive_dataframe_e %>%
  mutate(predicted_win = ifelse(prob_player1_wins > 0.5, 1, 0)) %>%
  summarize(accuracy = mean(predicted_win == outcome))

print(correct_predictions)
```
```{r}
dim(predictive_dataframe_e)
```

Calculate log loss:

```{r}
log_loss <- predictive_dataframe_e %>%
  mutate(log_loss = - (outcome * log(prob_player1_wins) + (1 - outcome) * log(1 - prob_player1_wins))) %>%
  summarize(mean_log_loss = mean(log_loss))

print(log_loss)
```
Calculating Brier Score:

```{r}
brier_score <- predictive_dataframe_e %>%
  mutate(brier_score = (prob_player1_wins - outcome)^2) %>%
  summarize(mean_brier_score = mean(brier_score))

print(brier_score)
```


## 8 COMPARATIVE ANALYSIS

Merging odds and predictive:

```{r}
# merge the odds_data with the original predictive_dataframe
merged_original <- predictive_dataframe_e %>%
  left_join(odds_data, by = c("EP_game_id", "player1", "player2")) %>%
  rename(
    golfer1_win_orig = golfer1_win,
    golfer2_win_orig = golfer2_win
  )


# swap player1 and player2 in odds_data for the merge
odds_data_swapped <- odds_data %>%
  rename(
    player1 = player2,
    player2 = player1,
    golfer1_win = golfer2_win,
    golfer2_win = golfer1_win
  )

# merge the odds_data with the predictive_dataframe with swapped order
merged_swapped <- predictive_dataframe_e %>%
  left_join(odds_data_swapped, by = c("EP_game_id", "player1", "player2")) %>%
  rename(
    golfer1_win_swapped = golfer1_win,
    golfer2_win_swapped = golfer2_win
  )


# combine the results from the original and swapped merges
predictive_odds_data_e <- merged_original %>%
  left_join(merged_swapped, by = c("EP_game_id", "player1", "player2", "outcome", "start_date", "prev_EP_game_id", "prev_tournament", "prev_major", "prev_start_date", "strength_player1", "strength_player2"))

# coalesce the columns to handle NA values and get the correct golfer1_win and golfer2_win
predictive_odds_data_e <- predictive_odds_data_e %>%
  mutate(
    tournament = coalesce(tournament.x.x, tournament.x.y, tournament.y.x, tournament.y.y),
    golfer1_win = coalesce(golfer1_win_orig, golfer1_win_swapped),
    golfer2_win = coalesce(golfer2_win_orig, golfer2_win_swapped),
    prob_player1_win = coalesce(prob_player1_wins.x, prob_player1_wins.y),
    prob_player2_win = coalesce(prob_player2_wins.x, prob_player2_wins.y),
    major = coalesce(major.x, major.y)
  ) %>%
  select(-golfer1_win_orig, -golfer2_win_orig, -golfer1_win_swapped, -golfer2_win_swapped, -tournament.x.x, -tournament.x.y, -tournament.y.x, -tournament.y.y, -prob_player1_wins.x, -prob_player1_wins.y, -prob_player2_wins.x, -prob_player2_wins.y,
         -major.x, -major.y)


```

Selecting relevant columns and deleting NA data:

```{r}
predictive_odds_data_e <- predictive_odds_data_e %>%
  select(EP_game_id, start_date, tournament, player1, player2, outcome ,prob_player1_win, prob_player2_win, golfer1_win, golfer2_win )

predictive_odds_data_e <- predictive_odds_data_e %>%
  drop_na()
```

Checking to make sure all is as it should be:

```{r}
dim(predictive_odds_data_e)
```

Calculating accuracy of these predictions:

```{r}
correct_predictions <- predictive_odds_data_e %>%
  mutate(predicted_win = ifelse(prob_player1_win > 0.5, 1, 0)) %>%
  summarize(accuracy = mean(predicted_win == outcome))

print(correct_predictions)
```
Calculate log loss:

```{r}
log_loss <- predictive_odds_data_e %>%
  mutate(log_loss = - (outcome * log(prob_player1_win) + (1 - outcome) * log(1 - prob_player1_win))) %>%
  summarize(mean_log_loss = mean(log_loss))

print(log_loss)
```
Calculating Brier Score:

```{r}
brier_score <- predictive_odds_data_e %>%
  mutate(brier_score = (prob_player1_win - outcome)^2) %>%
  summarize(mean_brier_score = mean(brier_score))

print(brier_score)
```

