---
title: "Standard Model"
output: html_document
date: "2024-07-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/lukeconkleton/Library/Mobile Documents/com~apple~CloudDocs/Documents/UCL Project/Data")

```


```{r, include = FALSE}
library(tidyverse)
library(BradleyTerry2)
library(elo)
library(igraph)
library(tidyr)
library(zoo)
library(forecast)
```

#############################################################################
#############################################################################

```{r}
rounds_data <- readRDS("roundData2018-2022.rds")
odds_data <- readRDS("oddsData2018-2022.rds")
```

## 1. DATA WRANGLING & FEATURE SELECTION

Subsetting dataframe to only include pre-covid data. Creating a new dataframe to only inlcude variables of interest; golfer, tournament, major, final position and date.

```{r}
rounds_data <- rounds_data %>% filter(year %in% c('2018', '2019'))
```

Engineering a feature for whether or not the golfer is playing at 'home':

```{r}
rounds_data$is.home <- ifelse(rounds_data$player_country == rounds_data$tournament_country, 1, 0)
```

```{r}
rounds <- data.frame(
  EP_game_id = rounds_data$EP_game_id,
  tournament = rounds_data$event_name,
  major = rounds_data$is_major,
  golfer_id = rounds_data$player_id,
  golfer = rounds_data$player_name,
  final_position = rounds_data$final_position,
  start_date = rounds_data$competition_start_date, 
  times_played_course = rounds_data$times_played_course,
  is.home = rounds_data$is.home
)

rounds <- rounds %>%
  distinct()
```

Only keeping a variable for times played course pre-tournament.

```{r}
rounds <- rounds %>%
  group_by(EP_game_id, tournament, golfer_id, start_date) %>%
  slice_min(times_played_course) %>% #takes the minimum value of the datapoints
  ungroup() %>%
  rename(times_played_course_pre_tournament = times_played_course)
```



Only considering the relevant time frame:

```{r}
odds_data$competition_start_date <- as.Date(odds_data$competition_start_date)

odds_data <- odds_data %>% filter(format(competition_start_date, "%Y") %in% c("2018", "2019"))

odds_data <- odds_data %>%
  select(EP_game_id, event_name, SMARTODDS_golfer1_name, SMARTODDS_golfer2_name, golfer1_win, golfer2_win) %>%
  rename(
    player1 = SMARTODDS_golfer1_name,
    player2 = SMARTODDS_golfer2_name,
    tournament = event_name
  )
```

More tournaments included in rounds data set comparatively:

```{r}
length(unique(odds_data$EP_game_id))
length(unique(rounds_data$EP_game_id))
```

Subsetting the rounds dataset to only include golfers who have made over 50 appearances in the odds dataset:

```{r}
odds_golfers <- c(odds_data$player1, odds_data$player2)
length(unique(odds_golfers))

# how many appearances in the odds data frame
golfer_counts <- data.frame(golfer = odds_golfers) %>%
  group_by(golfer) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

players_over_50 <- golfer_counts %>%
  filter(count >= 50)

dim(players_over_50)
```


```{r}
# Filter the original dataframe to keep only those players
filtered_rounds_data <- rounds_data %>%
  filter(player_name %in% players_over_50$golfer)

# Filter Rounds Dataframe

filtered_rounds <- rounds %>%
  filter(golfer %in% players_over_50$golfer) %>%
  filter(!is.na(final_position) & final_position != 0)
```

## 2. CREATING A PAIRWISE COMPARISON DATASET

Create a dataset with every single match-up in that time period. Can be modified to be grouped via 'EP_game_ID' or by 'tournament'. Similarly, is it easier to use indexes for the golfers or their names? For tournaments, need to sort by ID as there will be repetitions (two years means 2x each tournament).


```{r}
# Define the create_pairs function
create_pairs <- function(df) {
  pairs <- expand.grid(golfer_A = df$golfer_id, golfer_B = df$golfer_id)
  pairs <- pairs[pairs$golfer_A != pairs$golfer_B, ]
  
  golfer_A_indices <- match(pairs$golfer_A, df$golfer_id)
  golfer_B_indices <- match(pairs$golfer_B, df$golfer_id)
  
  if (length(golfer_A_indices) == 0 || length(golfer_B_indices) == 0) {
    return(data.frame()) 
  }
  
  result <- data.frame(
    golfer_A = df$golfer[golfer_A_indices],
    golfer_B = df$golfer[golfer_B_indices],
    final_position_A = df$final_position[golfer_A_indices],
    final_position_B = df$final_position[golfer_B_indices],
    start_date = df$start_date[golfer_A_indices],
    major = df$major[golfer_A_indices],
    EP_game_id = df$EP_game_id[golfer_A_indices],
    tournament = df$tournament[golfer_A_indices],
    playerA_times_played = df$times_played_course_pre_tournament[golfer_A_indices],
    playerB_times_played = df$times_played_course_pre_tournament[golfer_B_indices],
    is.home.A = df$is.home[golfer_A_indices],
    is.home.B = df$is.home[golfer_B_indices]
  )
  
  return(result)
}

# apply the function to each tournament
pairwise_matches <- filtered_rounds %>%
  group_by(EP_game_id) %>%
  do({
    result <- create_pairs(.)
    if (nrow(result) > 0) {
      result$EP_game_id <- .$EP_game_id[1]
    }
    result
  }) %>%
  ungroup()

pairwise_matches <- pairwise_matches %>%
  mutate(
    result_A = ifelse(final_position_A < final_position_B, 1, 0),
    result_B = ifelse(final_position_B < final_position_A, 1, 0)
  )

```

Ensure that there are no duplicate results in the dataset and that the data is ordered temporally. Will need to be amended if adding variables.

```{r}
# ensure each pair is in a consistent order
pairwise_matches <- pairwise_matches %>%
  rowwise() %>%
  mutate(
    golfer_A_sorted = ifelse(golfer_A < golfer_B, golfer_A, golfer_B),
    golfer_B_sorted = ifelse(golfer_A < golfer_B, golfer_B, golfer_A),
    final_position_A_sorted = ifelse(golfer_A < golfer_B, final_position_A, final_position_B),
    final_position_B_sorted = ifelse(golfer_A < golfer_B, final_position_B, final_position_A),
    result_A_sorted = ifelse(golfer_A < golfer_B, result_A, result_B),
    result_B_sorted = ifelse(golfer_A < golfer_B, result_B, result_A),
    is.home_A_sorted = ifelse(golfer_A < golfer_B, is.home.A, is.home.B),
    is.home_B_sorted = ifelse(golfer_A < golfer_B, is.home.B, is.home.A),
    times_played_A_sorted = ifelse(golfer_A < golfer_B, playerA_times_played, playerB_times_played),
    times_played_B_sorted = ifelse(golfer_A < golfer_B, playerB_times_played, playerA_times_played)
  ) %>%
  ungroup()

# Add home advantage columns after sorting
pairwise_matches <- pairwise_matches %>%
  mutate(
    home_advantage_A = ifelse(is.home_A_sorted == 1 & is.home_B_sorted == 0, 1, 0),
    home_advantage_B = ifelse(is.home_A_sorted == 0 & is.home_B_sorted == 1, 1, 0)
  )

# Remove duplicates based on the sorted pairs and the tournament
pairwise_matches <- pairwise_matches %>%
  select(EP_game_id, tournament, golfer_A_sorted, golfer_B_sorted, final_position_A_sorted, final_position_B_sorted, 
         result_A_sorted, result_B_sorted, start_date, major, times_played_A_sorted, times_played_B_sorted, 
         is.home_A_sorted, is.home_B_sorted, home_advantage_A, home_advantage_B) %>%
  distinct() %>%
  rename(
    golfer_A = golfer_A_sorted,
    golfer_B = golfer_B_sorted,
    final_position_A = final_position_A_sorted,
    final_position_B = final_position_B_sorted,
    result_A = result_A_sorted,
    result_B = result_B_sorted,
    is.home_A = is.home_A_sorted,
    is.home_B = is.home_B_sorted,
    times_played_A = times_played_A_sorted,
    times_played_B = times_played_B_sorted
  )

# Ensuring proper column naming for clarity
pairwise_matches <- pairwise_matches %>%
  rename(
    player1 = golfer_A,
    player2 = golfer_B,
    score1 = result_A,
    score2 = result_B
  )

# Maintaining temporal ordering
pairwise_matches <- pairwise_matches %>%
  arrange(start_date, EP_game_id)
```

#############################################################################
#############################################################################


## 3. CREATING A ROLLING WINDOW

Utilising a rolling window; takes a subset of the data and moves the window across the dataset with the window size constant(ish). The window slides forward by an observation at each step.

In this, an 'observation' is a tournamnet, so the window won't be necessarily have the same number of datapoints but will have the same number of tournaments.

This allows for the dynamic updating of the model as new data becomes available.

```{r}
unique_tournaments <- unique(pairwise_matches$EP_game_id)
length(unique_tournaments)
```

This function seems to be working.

```{r}
create_rolling_windows <- function(data, tournaments, window_size) {
  n <- length(tournaments)
  windows <- list()
  
  for (start in 1:(n - window_size + 1)) {
    end <- start + window_size - 1
    window_tournaments <- tournaments[start:end]
    window_data <- data[data$EP_game_id %in% window_tournaments, ]
    windows[[length(windows) + 1]] <- window_data
  }
  
  return(windows)
}

# window size (number of tournaments per window)
window_size <- 80

windows <- create_rolling_windows(pairwise_matches, unique_tournaments, window_size)

## make sure function is working as expected

length(windows)

unique(windows[[3]]$EP_game_id)
```

Create a window dataframe.

```{r}
create_window_summary <- function(windows) {
  summary_df <- data.frame(window_number = integer(),
                           final_EP_game_id = integer(),
                           final_tournament = character(),
                           start_data = character(),
                           major = integer(),
                           stringsAsFactors = FALSE)
  
  for (i in 1:(length(windows) - 1)) {
    next_window <- windows[[i + 1]]
    final_EP_game_id <- tail(next_window$EP_game_id, n = 1)  
    final_tournament <- tail(next_window$tournament, n = 1)  
    final_start_date <- tail(next_window$start_date, n = 1)  
    final_major <- tail(next_window$major, n = 1)  
    
    summary_df <- rbind(summary_df, data.frame(window_number = i,
                                               next_EP_game_id = final_EP_game_id,
                                               next_tournament = final_tournament,
                                               start_date = final_start_date,
                                               major = final_major))
  }
  
  return(summary_df)
}

window_summary <- create_window_summary(windows)

```

## 4. PREPARING DATA FOR TRAINING

Prepare the data so it is compatible with the Bradley-Terry package.

```{r}
prepare_pairwise_data <- function(data) {
  players <- unique(c(data$player1, data$player2))
  
  pairwise_data <- data.frame(
    player1 = factor(data$player1, levels = players),
    player2 = factor(data$player2, levels = players),
    wins1 = data$score1,
    wins2 = data$score2,
    major = data$major,
    played_before1 = data$times_played_A,
    played_before2 = data$times_played_B,
    is.home1 = data$is.home_A,
    is.home2 = data$is.home_B
  )
  
  return(pairwise_data)
}

```

```{r}
# Standardize factor levels function
standardize_levels <- function(data, all_players) {
  data$player1 <- factor(data$player1, levels = all_players)
  data$player2 <- factor(data$player2, levels = all_players)
  return(data)
}

# Collect all unique player names across all windows
all_players <- unique(unlist(lapply(windows, function(window) {
  unique(c(window$player1, window$player2))
})))

```

## 5. TRAINING MODEL

Train the B-T model for each individual window. At the moment only shows the first 5 windows. If including a reference player in the 'refcat' argument, their strength will be = 0 for every window.

```{r}
# Train Bradley-Terry model
train_bradley_terry <- function(window_data, window_number, ref_player) {
  pairwise_data <- prepare_pairwise_data(window_data)
  cat("Training model for window", window_number, "\n")
  model <- BTm(cbind(wins1, wins2), 
               player1 = player1, 
               player2 = player2, 
               data = pairwise_data,
               id = "player",
               refcat = ref_player)
  return(model)
}

reference_player <- "Tommy Fleetwood" # basis of comparison

# can choose how many windows to go through

models <- lapply(seq_len(min(100, length(windows))), function(i) {
  train_bradley_terry(windows[[i]], i, reference_player)
})


```

Create a dataframe to compare the coefficients from each window (model)

```{r}
# extract coefficients from each model and store them in a list of DataFrames
extract_coefficients <- function(models, window_numbers, ref_player) {
  coef_list <- lapply(window_numbers, function(i) {
    if (i <= length(models)) {
      model <- models[[i]]
      coefs <- coef(model)
      if (!(ref_player %in% names(coefs))) {
        coefs <- c(coefs, setNames(0, ref_player))
      }
      data.frame(
        Player = names(coefs),
        Coefficient = as.numeric(coefs),
        Window = paste("Window", i, sep = "_"),
        stringsAsFactors = FALSE
      )
    } else {
      data.frame(
        Player = character(0),
        Coefficient = numeric(0),
        Window = character(0),
        stringsAsFactors = FALSE
      )
    }
  })
  return(coef_list)
}


# Extract coefficients from models
window_numbers <- seq_len(length(models))
coef_list <- extract_coefficients(models, window_numbers, reference_player)

# Combine the list of DataFrames into a single DataFrame
combined_df <- bind_rows(coef_list)

# Ensure the reference player is in the final DataFrame
if (!("Tommy Fleetwood" %in% combined_df$Player)) {
  for (i in window_numbers) {
    combined_df <- combined_df %>%
      add_row(Player = reference_player, Coefficient = 0, Window = paste("Window", i, sep = "_"))
  }
}

# Reshape the DataFrame so each column represents a window
final_df <- combined_df %>%
  pivot_wider(names_from = Window, values_from = Coefficient) %>%
  arrange(Player)

# Remove leading dots from the 'Player' column in final_df
final_df$Player <- gsub("^player", "", final_df$Player)
```

Plot the coefficient values to assess whether these are respectable results.

Selecting certain golfers to visualise.

```{r}
selected_golfers <- c("Tommy Fleetwood", "Brooks Koepka", "Tony Finau", "Patrick Reed", "Bryson DeChambeau", "Rory McIlroy")

selected_df <- final_df %>%
  filter(Player %in% selected_golfers)

selected_plot_df <- selected_df %>%
  pivot_longer(cols = starts_with("Window_"), names_to = "Window", values_to = "Coefficient")

# Ensure Window is an ordered factor
selected_plot_df$Window <- factor(selected_plot_df$Window, levels = paste("Window", 1:length(windows), sep = "_"))

# Plot the coefficients for the selected golfers across windows
ggplot(selected_plot_df, aes(x = Window, y = Coefficient, group = Player, color = Player)) +
  geom_line() +
  geom_point() +
  theme(legend.position = "right") +  # Include legend for clarity
  labs(title = "Bradley-Terry Log-Strengths Across Windows for Selected Golfers", subtitle = "Reference Player: Tommy Fleetwood",
       x = "Window", y = "Coefficient") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(
    breaks = levels(selected_plot_df$Window)[seq(0, length(levels(selected_plot_df$Window)), by = 5)],
    labels = seq(5, length(levels(selected_plot_df$Window)), by = 5))
```


## PREDICTION ANALYSIS

```{r}
predictive_dataframe_s <- merge(pairwise_matches, window_summary, 
                       by.x = "EP_game_id", by.y = "next_EP_game_id", all.x = TRUE) %>%
  select(window_number, EP_game_id, player1, player2, tournament, score1, score2, major.x) %>%
  rename(player1_win = score1, major = major.x)
```

Now merging strengths:

```{r}
# merging the strengths 
predictive_dataframe_s <- predictive_dataframe_s %>%
  left_join(final_df %>% 
              gather(Window, Strength, -Player) %>% 
              mutate(Window = as.numeric(gsub("Window_", "", Window))), 
            by = c("window_number" = "Window", "player1" = "Player")) %>%
  rename(player1_strength = Strength) %>%
  left_join(final_df %>% 
              gather(Window, Strength, -Player) %>% 
              mutate(Window = as.numeric(gsub("Window_", "", Window))), 
            by = c("window_number" = "Window", "player2" = "Player")) %>%
  rename(player2_strength = Strength)
```

Adding column for model probabilities. The output of the package produces log strengths of players so probabilities have to be calculated accordingly:

```{r}
logistic <- function(x){
  1 / (1+exp(-x))
}

predictive_dataframe_s <- predictive_dataframe_s %>%
  mutate(
    prob_player1_win = logistic(player1_strength - player2_strength),
    prob_player2_win = logistic(player2_strength - player1_strength)
  )
```

Calculating Accuracy:

```{r}
# get rid of all non-predictive data
predictive_dataframe_s <- predictive_dataframe_s %>%
  drop_na()

correct_predictions <- predictive_dataframe_s %>%
  mutate(predicted_win = ifelse(prob_player1_win > 0.5, 1, 0)) %>%
  summarize(
    accuracy = mean(predicted_win == player1_win),
    accuracy_sd = sd(predicted_win == player1_win),  # Standard deviation of accuracy
    n = n(),  # Sample size
    accuracy_se = accuracy_sd / sqrt(n),  # Standard error
    ci_lower = accuracy - qnorm(0.975) * accuracy_se,  # Lower bound of 95% CI
    ci_upper = accuracy + qnorm(0.975) * accuracy_se   # Upper bound of 95% CI
  )

# Print the results
print(correct_predictions)
```

```{r}
dim(predictive_dataframe_s)
```

Calculating log loss:

```{r}
log_loss <- predictive_dataframe_s %>%
  mutate(log_loss = - (player1_win * log(prob_player1_win) + (1 - player1_win) * log(1 - prob_player1_win))) %>%
  summarize(mean_log_loss = mean(log_loss))

print(log_loss)
```
Calculating Brier Score:

```{r}
brier_score <- predictive_dataframe_s %>%
  mutate(brier_score = (prob_player1_win - player1_win)^2) %>%
  summarize(mean_brier_score = mean(brier_score))

print(brier_score)
```

Creating a calibration plot:

```{r}
# Create bins for predicted probabilities
predictive_dataframe_s <- predictive_dataframe_s %>%
  mutate(prob_bin = cut(prob_player1_win, breaks = seq(0, 1, by = 0.2), include.lowest = TRUE))

# Calculate the observed win rate in each bin
calibration_data <- predictive_dataframe_s %>%
  group_by(prob_bin) %>%
  summarize(observed_win_rate = mean(player1_win),
            predicted_win_rate = mean(prob_player1_win),
            count = n())

# Plot the calibration plot
ggplot(calibration_data, aes(x = predicted_win_rate, y = observed_win_rate)) +
  geom_point() +
  geom_line() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(title = "Calibration Plot", x = "Predicted Win Rate", y = "Observed Win Rate")
```


##8. PROFITABILITY ANALYSIS

Merging the odds dataframe with the predictive dataframe:

```{r}
# Step 1: Merge with Original Order
# Merge the odds_data with the original predictive_dataframe
merged_original <- predictive_dataframe_s %>%
  left_join(odds_data, by = c("EP_game_id", "player1", "player2")) %>%
  rename(
    golfer1_win_orig = golfer1_win,
    golfer2_win_orig = golfer2_win
  )

# Step 2: Merge with Swapped Order
# Swap player1 and player2 in odds_data for the merge
odds_data_swapped <- odds_data %>%
  rename(
    player1 = player2,
    player2 = player1,
    golfer1_win = golfer2_win,
    golfer2_win = golfer1_win
  )

# Merge the odds_data with the predictive_dataframe with swapped order
merged_swapped <- predictive_dataframe_s %>%
  left_join(odds_data_swapped, by = c("EP_game_id", "player1", "player2")) %>%
  rename(
    golfer1_win_swapped = golfer1_win,
    golfer2_win_swapped = golfer2_win
  )

# Step 3: Combine the Results
# Combine the results from the original and swapped merges
predictive_odds_data_s <- merged_original %>%
  left_join(merged_swapped, by = c("EP_game_id", "player1", "player2", "player1_win",   "player1_strength", "player2_strength", "major"))

# Coalesce the columns to handle NA values and get the correct golfer1_win and golfer2_win
predictive_odds_data_s <- predictive_odds_data_s %>%
  mutate(
    tournament = coalesce(tournament.x.x, tournament.x.y, tournament.y.x, tournament.y.y),
    golfer1_win = coalesce(golfer1_win_orig, golfer1_win_swapped),
    golfer2_win = coalesce(golfer2_win_orig, golfer2_win_swapped),
    prob_player1_win = coalesce(prob_player1_win.x, prob_player1_win.y),
    prob_player2_win = coalesce(prob_player2_win.x, prob_player2_win.y)
  ) %>%
  select(-golfer1_win_orig, -golfer2_win_orig, -golfer1_win_swapped, -golfer2_win_swapped, -tournament.x.x, -tournament.x.y, -tournament.y.x, -tournament.y.y, -prob_player1_win.x, -prob_player1_win.y, -prob_player2_win.x, -prob_player2_win.y,  window_number.x, major) %>%
  rename(window_number = window_number.x)
```

Selecting relevant columns and deleting data used in the first window:

```{r}
predictive_odds_data_s <- predictive_odds_data_s %>%
  select(EP_game_id, tournament, player1, player2, golfer1_win, golfer2_win, prob_player1_win, prob_player2_win, player1_win, window_number, major)

predictive_odds_data_s <- predictive_odds_data_s %>%
  drop_na()
```

Calculating dimensions:

```{r}
dim(predictive_odds_data_s)
```


Calculating Accuracy:

```{r}
correct_predictions <- predictive_odds_data_s %>%
  mutate(predicted_win = ifelse(prob_player1_win > 0.5, 1, 0)) %>%
  summarize(accuracy = mean(predicted_win == player1_win))

print(correct_predictions)
```
Calculate log loss:

```{r}
log_loss <- predictive_odds_data_s %>%
  mutate(log_loss = - (player1_win * log(prob_player1_win) + (1 - player1_win) * log(1 - prob_player1_win))) %>%
  summarize(mean_log_loss = mean(log_loss))

print(log_loss)
```
Calculating Brier Score:

```{r}
brier_score <- predictive_odds_data_s %>%
  mutate(brier_score = (prob_player1_win - player1_win)^2) %>%
  summarize(mean_brier_score = mean(brier_score))

print(brier_score)
```

Formulating a betting strategy. 
1. Calculate implied probabilities from Odds (1/odds)
2. Identify the value bets (prob greater than implied)
3. Two stake strategies, Kelly and unit (calculate for both)
  - Total Stake, Total pnL RoI
  - Total EV RoI

Adding a column with implied probabilities:

```{r}
predictive_odds_data_s <- predictive_odds_data_s %>%
  mutate(
    implied_prob_player1 = 1 / golfer1_win,
    implied_prob_player2 = 1 / golfer2_win
  )
```

```{r}
# Filter the data frame to only include rows where both final columns are FALSE
filtered_df <- predictive_odds_data_s[predictive_odds_data_s$value_bet_player1 == FALSE & predictive_odds_data_s$value_bet_player2 == FALSE, ]
```


Identify the value bets where model's predicted probability is greater than the bookmaker's implied probability:

```{r}
predictive_odds_data_s <- predictive_odds_data_s %>%
  mutate(
    value_bet_player1 = (prob_player1_win - implied_prob_player1) > 0,
    value_bet_player2 = (prob_player2_win - implied_prob_player2) > 0 
  )

# filtering out the dataframe for which both value_bets are FALSE:

predictive_odds_data_s <- predictive_odds_data_s %>%
  filter(value_bet_player1 == TRUE | value_bet_player2 == TRUE)
```

Creating two dataframes for stake strategy comparisons:

```{r}
predictive_odds_data_s <- predictive_odds_data_s %>%
  mutate(
    b_player1 = golfer1_win -1,
    b_player2 = golfer2_win -1, 
         
    q_player1 = 1 - prob_player1_win,
    q_player2 = 1 - prob_player2_win,
    
    kelly_fraction_player1 = ifelse(value_bet_player1, 
                                 (b_player1 * prob_player1_win - q_player1) / b_player1, 
                                 0),
    kelly_fraction_player2 = ifelse(value_bet_player2, 
                                 (b_player2 * prob_player2_win - q_player2) / b_player2, 
                                 0)) %>%
  select(-b_player1, -b_player2, - q_player1, -q_player2)
```


Implementing 3 Strategies:


```{r}
bankroll_kelly <- 1000
bankroll_half_kelly <- 1000
bankroll_unit <- 1000

# Start with the initial bankroll, no stakes, no profit/loss, and no expected value
bankroll_kelly_progress <- c(bankroll_kelly)
stake_kelly_progress <- c(0)
profit_loss_kelly_progress <- c(0)
expected_value_kelly_progress <- c(0)

bankroll_half_kelly_progress <- c(bankroll_half_kelly)
stake_half_kelly_progress <- c(0)
profit_loss_half_kelly_progress <- c(0)
expected_value_half_kelly_progress <- c(0)

bankroll_unit_progress <- c(bankroll_unit)
stake_unit_progress <- c(0)
profit_loss_unit_progress <- c(0)
expected_value_unit_progress <- c(0)

# Add an initial window 0 to represent the starting state
windows <- c(0, sort(unique(predictive_odds_data_s$window_number)))
```


```{r}
for (window in windows[-1]) {  # Skip the first window 0 as it's already initialized
    # Filter bets for the current window
    bets_in_window <- subset(predictive_odds_data_s, window_number == window)
    
    # Calculate total Kelly fractions for normalization if necessary
    total_fraction <- sum(bets_in_window$kelly_fraction_player1) + sum(bets_in_window$kelly_fraction_player2)
    if (total_fraction > 1) {
        normalization_factor <- 1 / total_fraction
        bets_in_window$kelly_fraction_player1 <- bets_in_window$kelly_fraction_player1 * normalization_factor
        bets_in_window$kelly_fraction_player2 <- bets_in_window$kelly_fraction_player2 * normalization_factor
    } else {
        normalization_factor <- 1
    }

    # Reset stakes, profits/losses, and expected value for the current window
    stake_kelly_window <- 0
    profit_loss_kelly_window <- 0
    expected_value_kelly_window <- 0
    
    stake_half_kelly_window <- 0
    profit_loss_half_kelly_window <- 0
    expected_value_half_kelly_window <- 0
    
    stake_unit_window <- 0
    profit_loss_unit_window <- 0
    expected_value_unit_window <- 0
    
    # Kelly Stake
    for (i in 1:nrow(bets_in_window)) {
        bet <- bets_in_window[i, ]
        
        if (bet$value_bet_player1) {
            stake_kelly <- bankroll_kelly * bet$kelly_fraction_player1
            stake_kelly_window <- stake_kelly_window + stake_kelly
            
            expected_value_kelly <- stake_kelly * bet$golfer1_win - stake_kelly * (1 - bet$golfer1_win)
            expected_value_kelly_window <- expected_value_kelly_window + expected_value_kelly
            
            if (bet$player1_win == 1) {
                profit_loss_kelly_window <- profit_loss_kelly_window + stake_kelly * (bet$golfer1_win - 1)
            } else {
                profit_loss_kelly_window <- profit_loss_kelly_window - stake_kelly
            }
        }
        
        if (bet$value_bet_player2) {
            stake_kelly <- bankroll_kelly * bet$kelly_fraction_player2
            stake_kelly_window <- stake_kelly_window + stake_kelly
            
            expected_value_kelly <- stake_kelly * bet$golfer2_win - stake_kelly * (1 - bet$golfer2_win)
            expected_value_kelly_window <- expected_value_kelly_window + expected_value_kelly
            
            if (bet$player1_win == 0) {
                profit_loss_kelly_window <- profit_loss_kelly_window + stake_kelly * (bet$golfer2_win - 1)
            } else {
                profit_loss_kelly_window <- profit_loss_kelly_window - stake_kelly
            }
        }
    }
    
    # Update bankroll after the window
    bankroll_kelly <- bankroll_kelly + profit_loss_kelly_window

    # Store the results for Kelly Stake
    bankroll_kelly_progress <- c(bankroll_kelly_progress, bankroll_kelly)
    stake_kelly_progress <- c(stake_kelly_progress, stake_kelly_window)
    profit_loss_kelly_progress <- c(profit_loss_kelly_progress, profit_loss_kelly_window)
    expected_value_kelly_progress <- c(expected_value_kelly_progress, expected_value_kelly_window)
    
    # Repeat similar steps for Half Kelly Stake and Unit Stake
    # Half Kelly Stake
    for (i in 1:nrow(bets_in_window)) {
        bet <- bets_in_window[i, ]
        
        if (bet$value_bet_player1) {
            stake_half_kelly <- bankroll_half_kelly * (bet$kelly_fraction_player1 / 2)
            stake_half_kelly_window <- stake_half_kelly_window + stake_half_kelly
            
            expected_value_half_kelly <- stake_half_kelly * bet$golfer1_win - stake_half_kelly * (1 - bet$golfer1_win)
            expected_value_half_kelly_window <- expected_value_half_kelly_window + expected_value_half_kelly
            
            if (bet$player1_win == 1) {
                profit_loss_half_kelly_window <- profit_loss_half_kelly_window + stake_half_kelly * (bet$golfer1_win - 1)
            } else {
                profit_loss_half_kelly_window <- profit_loss_half_kelly_window - stake_half_kelly
            }
        }
        
        if (bet$value_bet_player2) {
            stake_half_kelly <- bankroll_half_kelly * (bet$kelly_fraction_player2 / 2)
            stake_half_kelly_window <- stake_half_kelly_window + stake_half_kelly
            
            expected_value_half_kelly <- stake_half_kelly * bet$golfer2_win - stake_half_kelly * (1 - bet$golfer2_win)
            expected_value_half_kelly_window <- expected_value_half_kelly_window + expected_value_half_kelly
            
            if (bet$player1_win == 0) {
                profit_loss_half_kelly_window <- profit_loss_half_kelly_window + stake_half_kelly * (bet$golfer2_win - 1)
            } else {
                profit_loss_half_kelly_window <- profit_loss_half_kelly_window - stake_half_kelly
            }
        }
    }
    
    bankroll_half_kelly <- bankroll_half_kelly + profit_loss_half_kelly_window
    
    bankroll_half_kelly_progress <- c(bankroll_half_kelly_progress, bankroll_half_kelly)
    stake_half_kelly_progress <- c(stake_half_kelly_progress, stake_half_kelly_window)
    profit_loss_half_kelly_progress <- c(profit_loss_half_kelly_progress, profit_loss_half_kelly_window)
    expected_value_half_kelly_progress <- c(expected_value_half_kelly_progress, expected_value_half_kelly_window)
    
    # Unit Stake
    total_unit_fraction <- nrow(bets_in_window) * 0.05
    if (total_unit_fraction > 1) {
        normalization_factor <- 1 / total_unit_fraction
    } else {
        normalization_factor <- 1
    }
    
    for (i in 1:nrow(bets_in_window)) {
        bet <- bets_in_window[i, ]
        stake_unit <- bankroll_unit * 0.05 * normalization_factor
        stake_unit_window <- stake_unit_window + stake_unit
        
        expected_value_unit <- stake_unit * bet$golfer1_win - stake_unit * (1 - bet$golfer1_win)
        expected_value_unit_window <- expected_value_unit_window + expected_value_unit
        
        if (bet$value_bet_player1) {
            if (bet$player1_win == 1) {
                profit_loss_unit_window <- profit_loss_unit_window + stake_unit * (bet$golfer1_win - 1)
            } else {
                profit_loss_unit_window <- profit_loss_unit_window - stake_unit
            }
        }
        
        if (bet$value_bet_player2) {
            if (bet$player1_win == 0) {
                profit_loss_unit_window <- profit_loss_unit_window + stake_unit * (bet$golfer2_win - 1)
            } else {
                profit_loss_unit_window <- profit_loss_unit_window - stake_unit
            }
        }
    }
    
    bankroll_unit <- bankroll_unit + profit_loss_unit_window
    
    bankroll_unit_progress <- c(bankroll_unit_progress, bankroll_unit)
    stake_unit_progress <- c(stake_unit_progress, stake_unit_window)
    profit_loss_unit_progress <- c(profit_loss_unit_progress, profit_loss_unit_window)
    expected_value_unit_progress <- c(expected_value_unit_progress, expected_value_unit_window)
}
```

Datframe for all the important metrics:

```{r}
results_s <- data.frame(
    window = windows,
    bankroll_kelly = bankroll_kelly_progress,
    bankroll_half_kelly = bankroll_half_kelly_progress,
    bankroll_unit = bankroll_unit_progress,
    stake_kelly = stake_kelly_progress,
    stake_half_kelly = stake_half_kelly_progress,
    stake_unit = stake_unit_progress,
    profit_loss_kelly = profit_loss_kelly_progress,
    profit_loss_half_kelly = profit_loss_half_kelly_progress,
    profit_loss_unit = profit_loss_unit_progress,
    expected_value_kelly = expected_value_kelly_progress,
    expected_value_half_kelly = expected_value_half_kelly_progress,
    expected_value_unit = expected_value_unit_progress
)

# Calculate EV ROI
results_s$ev_roi_kelly <- results_s$expected_value_kelly / results_s$stake_kelly * 100
results_s$ev_roi_half_kelly <- results_s$expected_value_half_kelly / results_s$stake_half_kelly * 100
results_s$ev_roi_unit <- results_s$expected_value_unit / results_s$stake_unit * 100

# Calculate Actual ROI
results_s$roi_kelly <- results_s$profit_loss_kelly / results_s$stake_kelly * 100
results_s$roi_half_kelly <- results_s$profit_loss_half_kelly / results_s$stake_half_kelly * 100
results_s$roi_unit <- results_s$profit_loss_unit / results_s$stake_unit * 100

results_s[is.na(results_s) | results_s == Inf] <- 0
```

Actual ROI vs Expected ROI:

```{r}
filtered_results_s <- results_s %>% filter(window >= 1)

# Plotting the filtered data
ggplot(filtered_results_s, aes(x = window)) +
  geom_line(aes(y = roi_kelly, color = "Kelly Stake", linetype = "Actual")) +
  geom_line(aes(y = ev_roi_kelly, color = "Kelly Stake", linetype = "Expected")) +
  geom_line(aes(y = roi_unit, color = "Unit Stake", linetype = "Actual")) +
  geom_line(aes(y = ev_roi_unit, color = "Unit Stake", linetype = "Expected")) +
  scale_color_manual(values = c("Kelly Stake" = "blue", "Unit Stake" = "red")) + # Custom colors
  labs(title = "Actual vs. Expected ROI Over Time",
       x = "Window",
       y = "ROI (%)",
       color = "Staking Strategy",
       linetype = "ROI Type") +
  theme_minimal()
```

Cumulative Profit / Loss:

```{r}
results_s$cumulative_profit_loss_kelly <- cumsum(results_s$profit_loss_kelly)
results_s$cumulative_profit_loss_half_kelly <- cumsum(results_s$profit_loss_half_kelly)
results_s$cumulative_profit_loss_unit <- cumsum(results_s$profit_loss_unit)

ggplot(results_s, aes(x = window)) +
    geom_line(aes(y = cumulative_profit_loss_kelly, color = "Kelly Stake")) +
    geom_line(aes(y = cumulative_profit_loss_half_kelly, color = "Half Kelly Stake")) +
    geom_line(aes(y = cumulative_profit_loss_unit, color = "Unit Stake")) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "steelblue") +
    labs(title = "Cumulative Profit/Loss Over Time",
         x = "Window",
         y = "Cumulative Profit/Loss",
         color = "Staking Strategy") +
    scale_color_manual(values = c("Kelly Stake" = "blue", "Half Kelly Stake" = "red", "Unit Stake" = "green")) +
    theme_minimal()
```

Heatmap of ROI across windows:

```{r}
results_melted <- reshape2::melt(results_s, id.vars = "window", measure.vars = c("roi_kelly", "roi_unit"))

ggplot(results_melted, aes(x = window, y = variable, fill = value)) +
    geom_tile() +
    scale_fill_gradient(low = "blue", high = "red") +
    labs(title = "Heatmap of ROI Across Windows",
         x = "Window",
         y = "Staking Strategy",
         fill = "ROI (%)") +
    theme_minimal()
```

Boxplot of ROI across:

```{r}
results_melted <- reshape2::melt(results_s, id.vars = "window", measure.vars = c("roi_kelly", "roi_unit"))

ggplot(results_melted, aes(x = variable, y = value, fill = variable)) +
    geom_boxplot() +
    labs(title = "Distribution of ROI for Each Strategy",
         x = "Staking Strategy",
         y = "ROI (%)",
         fill = "Staking Strategy") +
    theme_minimal()
```





