---
title: "PageRank"
output: html_document
date: "2024-07-28"
---

```{r, include=FALSE}
library(tidyverse)
library(igraph)
```

#############################################################################
#############################################################################

```{r}
rounds_data <- readRDS("roundData2018-2022.rds")
odds_data <- readRDS("oddsData2018-2022.rds")
```

## 1. DATA WRANGLING & FEATURE SELECTION

Subsetting dataframe to only include pre-covid data. Creating a new dataframe to only inlcude variables of interest; golfer, tournament, major, final position and date.

```{r}
rounds_data <- rounds_data %>% filter(year %in% c('2018', '2019'))
```

Engineering a feature for whether or not the golfer is playing at 'home':

```{r}
rounds_data$is.home <- ifelse(rounds_data$player_country == rounds_data$tournament_country, 1, 0)
```

```{r}
rounds <- data.frame(
  EP_game_id = rounds_data$EP_game_id,
  tournament = rounds_data$event_name,
  major = rounds_data$is_major,
  golfer_id = rounds_data$player_id,
  golfer = rounds_data$player_name,
  final_position = rounds_data$final_position,
  start_date = rounds_data$competition_start_date, 
  times_played_course = rounds_data$times_played_course,
  is.home = rounds_data$is.home
)

rounds <- rounds %>%
  distinct()
```

Only keeping a variable for times played course pre-tournament.

```{r}
rounds <- rounds %>%
  group_by(EP_game_id, tournament, golfer_id, start_date) %>%
  slice_min(times_played_course) %>% #takes the minimum value of the datapoints
  ungroup() %>%
  rename(times_played_course_pre_tournament = times_played_course)
```



Only considering the relevant time frame:

```{r}
odds_data$competition_start_date <- as.Date(odds_data$competition_start_date)

odds_data <- odds_data %>% filter(format(competition_start_date, "%Y") %in% c("2018", "2019"))

odds_data <- odds_data %>%
  select(EP_game_id, event_name, SMARTODDS_golfer1_name, SMARTODDS_golfer2_name, golfer1_win, golfer2_win) %>%
  rename(
    player1 = SMARTODDS_golfer1_name,
    player2 = SMARTODDS_golfer2_name,
    tournament = event_name
  )
```

More tournaments included in rounds data set comparatively:

```{r}
length(unique(odds_data$EP_game_id))
length(unique(rounds_data$EP_game_id))
```

How many times does each golfer appear in the odds dataframe?

a)  Subsetting the rounds dataset to only include golfers who have made over 'x' appearances in the odds dataset:

```{r}
odds_golfers <- c(odds_data$player1, odds_data$player2)
length(unique(odds_golfers))

# how many appearances in the odds data frame
golfer_counts <- data.frame(golfer = odds_golfers) %>%
  group_by(golfer) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

players_over_50 <- golfer_counts %>%
  filter(count >= 50)

dim(players_over_50)
```


```{r}
# filter the original dataframe to keep only those players
filtered_rounds_data <- rounds_data %>%
  filter(player_name %in% players_over_50$golfer)

# filter Rounds Dataframe

filtered_rounds <- rounds %>%
  filter(golfer %in% players_over_50$golfer) %>%
  filter(!is.na(final_position) & final_position != 0)
```

## 2. CREATING A PAIRWISE COMPARISON DATASET

Create a dataset with every single match-up in that time period. 


```{r}

create_pairs <- function(df) {
  pairs <- expand.grid(golfer_A = df$golfer_id, golfer_B = df$golfer_id)
  pairs <- pairs[pairs$golfer_A != pairs$golfer_B, ]
  
  golfer_A_indices <- match(pairs$golfer_A, df$golfer_id)
  golfer_B_indices <- match(pairs$golfer_B, df$golfer_id)
  
  if (length(golfer_A_indices) == 0 || length(golfer_B_indices) == 0) {
    return(data.frame()) 
  }
  
  result <- data.frame(
    golfer_A = df$golfer[golfer_A_indices],
    golfer_B = df$golfer[golfer_B_indices],
    final_position_A = df$final_position[golfer_A_indices],
    final_position_B = df$final_position[golfer_B_indices],
    start_date = df$start_date[golfer_A_indices],
    major = df$major[golfer_A_indices],
    EP_game_id = df$EP_game_id[golfer_A_indices],
    tournament = df$tournament[golfer_A_indices],
    playerA_times_played = df$times_played_course_pre_tournament[golfer_A_indices],
    playerB_times_played = df$times_played_course_pre_tournament[golfer_B_indices],
    is.home.A = df$is.home[golfer_A_indices],
    is.home.B = df$is.home[golfer_B_indices]
  )
  
  return(result)
}

# apply the function to each tournament
pairwise_matches <- filtered_rounds %>%
  group_by(EP_game_id) %>%
  do({
    result <- create_pairs(.)
    if (nrow(result) > 0) {
      result$EP_game_id <- .$EP_game_id[1]
    }
    result
  }) %>%
  ungroup()

pairwise_matches <- pairwise_matches %>%
  mutate(
    result_A = ifelse(final_position_A < final_position_B, 1, 0),
    result_B = ifelse(final_position_B < final_position_A, 1, 0)
  )

```

Ensure that there are no duplicate results in the dataset and that the data is ordered temporally. Will need to be amended if adding variables.

```{r}
# ensure each pair is in a consistent order
pairwise_matches <- pairwise_matches %>%
  rowwise() %>%
  mutate(
    golfer_A_sorted = ifelse(golfer_A < golfer_B, golfer_A, golfer_B),
    golfer_B_sorted = ifelse(golfer_A < golfer_B, golfer_B, golfer_A),
    final_position_A_sorted = ifelse(golfer_A < golfer_B, final_position_A, final_position_B),
    final_position_B_sorted = ifelse(golfer_A < golfer_B, final_position_B, final_position_A),
    result_A_sorted = ifelse(golfer_A < golfer_B, result_A, result_B),
    result_B_sorted = ifelse(golfer_A < golfer_B, result_B, result_A),
    is.home_A_sorted = ifelse(golfer_A < golfer_B, is.home.A, is.home.B),
    is.home_B_sorted = ifelse(golfer_A < golfer_B, is.home.B, is.home.A),
    times_played_A_sorted = ifelse(golfer_A < golfer_B, playerA_times_played, playerB_times_played),
    times_played_B_sorted = ifelse(golfer_A < golfer_B, playerB_times_played, playerA_times_played)
  ) %>%
  ungroup()

# add home advantage columns after sorting
pairwise_matches <- pairwise_matches %>%
  mutate(
    home_advantage_A = ifelse(is.home_A_sorted == 1 & is.home_B_sorted == 0, 1, 0),
    home_advantage_B = ifelse(is.home_A_sorted == 0 & is.home_B_sorted == 1, 1, 0)
  )

# remove duplicates based on the sorted pairs and the tournament
pairwise_matches <- pairwise_matches %>%
  select(EP_game_id, tournament, golfer_A_sorted, golfer_B_sorted, final_position_A_sorted, final_position_B_sorted, 
         result_A_sorted, result_B_sorted, start_date, major, times_played_A_sorted, times_played_B_sorted, 
         is.home_A_sorted, is.home_B_sorted, home_advantage_A, home_advantage_B) %>%
  distinct() %>%
  rename(
    golfer_A = golfer_A_sorted,
    golfer_B = golfer_B_sorted,
    final_position_A = final_position_A_sorted,
    final_position_B = final_position_B_sorted,
    result_A = result_A_sorted,
    result_B = result_B_sorted,
    is.home_A = is.home_A_sorted,
    is.home_B = is.home_B_sorted,
    times_played_A = times_played_A_sorted,
    times_played_B = times_played_B_sorted
  )

# ensuring proper column naming for clarity
pairwise_matches <- pairwise_matches %>%
  rename(
    player1 = golfer_A,
    player2 = golfer_B,
    score1 = result_A,
    score2 = result_B
  )

# maintaining temporal ordering
pairwise_matches <- pairwise_matches %>%
  arrange(start_date, EP_game_id)
```

#############################################################################
#############################################################################

## 3 CREATING A ROLLING WINDOW

Utilising a rolling window; takes a subset of the data and moves the window across the dataset with the window size constant(ish). The window slides forward by an observation at each step.

In this, an 'observation' is a tournament, so the window won't be necessarily have the same number of datapoints but will have the same number of tournaments.

This allows for the dynamic updating of the model as new data becomes available.

```{r}
unique_tournaments <- unique(pairwise_matches$EP_game_id)
length(unique_tournaments)
```

This function seems to be working.

```{r}
create_rolling_windows <- function(data, tournaments, window_size) {
  n <- length(tournaments)
  windows <- list()
  
  for (start in 1:(n - window_size + 1)) {
    end <- start + window_size - 1
    window_tournaments <- tournaments[start:end]
    window_data <- data[data$EP_game_id %in% window_tournaments, ]
    windows[[length(windows) + 1]] <- window_data
  }
  
  return(windows)
}

# window size (number of tournaments per window)
window_size <- 80

windows <- create_rolling_windows(pairwise_matches, unique_tournaments, window_size)

## make sure function is working as expected

length(windows)
length(unique(windows[[4]]$EP_game_id))

```

Create a window dataframe.

```{r}
create_window_summary <- function(windows) {
  summary_df <- data.frame(window_number = integer(),
                           final_EP_game_id = integer(),
                           final_tournament = character(),
                           start_data = character(),
                           major = integer(),
                           stringsAsFactors = FALSE)
  
  for (i in 1:(length(windows) - 1)) {
    next_window <- windows[[i + 1]]
    final_EP_game_id <- tail(next_window$EP_game_id, n = 1)  
    final_tournament <- tail(next_window$tournament, n = 1)  
    final_start_date <- tail(next_window$start_date, n = 1)  
    final_major <- tail(next_window$major, n = 1)  
    
    summary_df <- rbind(summary_df, data.frame(window_number = i,
                                               next_EP_game_id = final_EP_game_id,
                                               next_tournament = final_tournament,
                                               start_date = final_start_date,
                                               major = final_major))
  }
  
  return(summary_df)
}

window_summary <- create_window_summary(windows)
```

## 4 PREPARING DATA FOR MODEL

## 5 TRAINING MODEL

Create an adjacency matrix:

```{r}
get_adjacency_matrix <- function(window_data) {
  players <- unique(c(window_data$player1, window_data$player2))
  n <- length(players)
  player_indices <- setNames(1:n, players)
  
  adj_matrix <- matrix(0, nrow = n, ncol = n)
  rownames(adj_matrix) <- players
  colnames(adj_matrix) <- players
  
  for (i in 1:nrow(window_data)) {
    if (window_data$score1[i] == 1) { # player1 won
      winner <- window_data$player1[i]
      loser <- window_data$player2[i]
    } else { # player2 won
      winner <- window_data$player2[i]
      loser <- window_data$player1[i]
    }
    
    adj_matrix[loser, winner] <- adj_matrix[loser, winner] + 1
  }
  
  return(adj_matrix)
}

```

Compute PageRank for each adjacency matrix:

```{r}
compute_pagerank <- function(adj_matrix) {
  g <- graph_from_adjacency_matrix(adj_matrix, mode = "directed", weighted = NULL)
  pr <- page_rank(g, damping = 0.85)$vector
  return(pr)
}
```

Get results:

```{r}
pagerank_results <- lapply(windows, function(window_data) {
  adj_matrix <- get_adjacency_matrix(window_data)
  pagerank <- compute_pagerank(adj_matrix)
  return(pagerank)
})

```

Making a plot of the directed graph (for window 10):

```{r}
selected_golfers <- c("Bubba Watson", "Erik van Rooyen", "Rory McIlroy")

# extract the adjacency matrix for the 10th window
adj_matrix_10 <- get_adjacency_matrix(windows[[10]])

golfer_indices <- which(rownames(adj_matrix_10) %in% selected_golfers)
adj_matrix_selected <- adj_matrix_10[golfer_indices, golfer_indices]

g_selected <- graph_from_adjacency_matrix(adj_matrix_selected, mode = "directed", weighted = NULL)

# plot the graph for the selected golfers
plot(g_selected, edge.arrow.size = 0.5, vertex.size = 30, 
     vertex.label.cex = 1.5, vertex.color = "lightblue", 
     edge.color = "gray", layout = layout_in_circle,
     main = "Directed Graph for Selected Golfers in 10th Window")

```

Creating a dataframe for player ranks:

```{r}
# get the unique list of players across all windows
all_players <- unique(unlist(lapply(windows, function(w) unique(c(w$player1, w$player2)))))

pagerank_df <- data.frame(matrix(NA, nrow = length(all_players), ncol = length(windows)))
rownames(pagerank_df) <- all_players
colnames(pagerank_df) <- paste0("Window_", seq_along(windows))

for (i in seq_along(pagerank_results)) {

  pr_scores <- pagerank_results[[i]]
  
  # Mmatch the players in the PageRank results to the rows in the dataframe
  pagerank_df[match(names(pr_scores), rownames(pagerank_df)), i] <- pr_scores
}


```

## PREDICTIVE ANALYSIS

Creating a predictive dataframe:

```{r}
predictive_dataframe_pr <- merge(pairwise_matches, window_summary, 
                       by.x = "EP_game_id", by.y = "next_EP_game_id", all.x = TRUE) %>%
  select(window_number, start_date.x, EP_game_id, tournament, player1, player2, final_position_A, final_position_B, score1, score2, times_played_A, times_played_B, is.home_A, is.home_B, major.x) %>%
  na.omit() %>% # gets rid of all data prior to window 1
  rename(major = major.x, start_date = start_date.x, player1_win = score1)
```

Formatting dataframe into long format so can add strengths:

```{r}
# assuming player names are rownames
pagerank_df <- pagerank_df %>% 
  rownames_to_column(var = "Player")

# reshape the data
pagerank_long <- pagerank_df %>%
  pivot_longer(cols = starts_with("Window_"),  # Use starts_with to gather all Window columns
               names_to = "Window", 
               values_to = "Rank")

# extract the window number from the 'Window' column
pagerank_long$Window <- as.numeric(sub("Window_", "", pagerank_long$Window))
```

Adding strengths:

```{r}
predictive_dataframe_pr <- predictive_dataframe_pr %>%
  left_join(pagerank_long %>% select(Player, Window, Rank), 
            by = c("player1" = "Player", "window_number" = "Window")) %>%
  rename(ranking_player1 = Rank) %>%
  left_join(pagerank_long %>% select(Player, Window, Rank), 
            by = c("player2" = "Player", "window_number" = "Window")) %>%
  rename(ranking_player2 = Rank)

```

Adding a column for predicted winners: P(A beats B) = rank_A / (rank_A + rank_B)

```{r}
predictive_dataframe_pr <- predictive_dataframe_pr %>%
  mutate(prob_player1_win = ranking_player1/(ranking_player1 + ranking_player2),
         prob_player2_win = ranking_player2/(ranking_player1 + ranking_player2))
```

```{r}
predictive_dataframe_pr <- predictive_dataframe_pr %>%
  select(-start_date, -final_position_A, -final_position_B, -times_played_A, -times_played_B, -is.home_A, -is.home_B, -score2)

```


Calculating Accuracy:

```{r}
# get rid of all non-predictive data
predictive_dataframe_pr <- predictive_dataframe_pr %>%
  drop_na()

correct_predictions <- predictive_dataframe_pr %>%
  mutate(predicted_win = ifelse(prob_player1_win > 0.5, 1, 0)) %>%
  summarize(accuracy = mean(predicted_win == player1_win))

print(correct_predictions)
```

```{r}
dim(predictive_dataframe_pr)
```


Calculating log loss:

```{r}
log_loss <- predictive_dataframe_pr %>%
  mutate(log_loss = - (player1_win * log(prob_player1_win) + (1 - player1_win) * log(1 - prob_player1_win))) %>%
  summarize(mean_log_loss = mean(log_loss))

print(log_loss)
```

Calculating Brier Score:

```{r}
brier_score <- predictive_dataframe_pr %>%
  mutate(brier_score = (prob_player1_win - player1_win)^2) %>%
  summarize(mean_brier_score = mean(brier_score))

print(brier_score)
```


Creating a calibration plot:

```{r}
# create bins for predicted probabilities
predictive_dataframe_pr <- predictive_dataframe_pr %>%
  mutate(prob_bin = cut(prob_player1_win, breaks = seq(0, 1, by = 0.2), include.lowest = TRUE))

# calculate the observed win rate in each bin
calibration_data <- predictive_dataframe_pr %>%
  group_by(prob_bin) %>%
  summarize(observed_win_rate = mean(player1_win),
            predicted_win_rate = mean(prob_player1_win),
            count = n())

ggplot(calibration_data, aes(x = predicted_win_rate, y = observed_win_rate)) +
  geom_point() +
  geom_line() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  labs(title = "Calibration Plot", x = "Predicted Win Rate", y = "Observed Win Rate")
```

Window Metrics:

```{r}
window_metrics_pr <- predictive_dataframe_pr %>%
    mutate(predicted_win = ifelse(prob_player1_win > 0.5, 1, 0)) %>%
    group_by(window_number, major) %>%
    summarize(
        accuracy = mean(predicted_win == player1_win),
        accuracy_sd = sd(predicted_win == player1_win, na.rm = TRUE),  # Calculate the standard deviation of accuracy
        num_matches = n(),
        mean_log_loss = mean(-(player1_win * log(prob_player1_win) + (1 - player1_win) * log(1 - prob_player1_win)), na.rm = TRUE),
        sd_log_loss = sd(-(player1_win * log(prob_player1_win) + (1 - player1_win) * log(1 - prob_player1_win)), na.rm = TRUE),  # Calculate the standard deviation of log-loss
        mean_brier_score = mean((prob_player1_win - player1_win)^2, na.rm = TRUE),
        sd_brier_score = sd((prob_player1_win - player1_win)^2, na.rm = TRUE)  # Calculate the standard deviation of Brier Score
    ) %>%
    filter(num_matches > 10)

ggplot(window_metrics_pr, aes(x = window_number)) +
  geom_line(aes(y = accuracy, color = "Accuracy")) +  # Accuracy line
  geom_point(data = subset(window_metrics_pr, major == 1), aes(y = accuracy, color = "Major"), size = 2) +  # Only Major points
  geom_line(aes(y = accuracy_sd, color = "Accuracy SD"), linetype = "dashed") +  # Rolling SD line
  geom_hline(yintercept = 0.5914805, linetype = "dotted", color = "black") +
  labs(
    title = "Accuracy and Rolling Standard Deviation of Accuracy per Window",
    x = "Window Number",
    y = "Value",
    color = "Metric"
  ) +
  scale_color_manual(values = c("Accuracy" = "navyblue", "Accuracy SD" = "orange", "Major" = "red")) +  # Custom colors including red for major
  theme_minimal()
```
```{r}
window_metrics_odds_pr <- predictive_odds_data_pr %>%
    mutate(predicted_win = ifelse(prob_player1_win > 0.5, 1, 0)) %>%
    group_by(window_number, major) %>%
    summarize(
        accuracy = mean(predicted_win == player1_win),
        accuracy_sd = sd(predicted_win == player1_win, na.rm = TRUE),
        num_matches = n(),
        mean_log_loss = mean(-(player1_win * log(prob_player1_win) + (1 - player1_win) * log(1 - prob_player1_win)), na.rm = TRUE),
        sd_log_loss = sd(-(player1_win * log(prob_player1_win) + (1 - player1_win) * log(1 - prob_player1_win)), na.rm = TRUE),  of log-loss
        mean_brier_score = mean((prob_player1_win - player1_win)^2, na.rm = TRUE),
        sd_brier_score = sd((prob_player1_win - player1_win)^2, na.rm = TRUE)  
    ) %>%
    filter(num_matches > 5)


ggplot(window_metrics_odds_pr, aes(x = window_number)) +
  geom_line(aes(y = accuracy, color = "Accuracy")) +  # Accuracy line
  geom_point(data = subset(window_metrics_odds_pr, major == 1), aes(y = accuracy, color = "Major"), size = 2) +  # Only Major points
  geom_line(aes(y = accuracy_sd, color = "Accuracy SD"), linetype = "dashed") +  # Rolling SD line
  geom_hline(yintercept = 0.5384615, linetype = "dotted", color = "black") +
  labs(
    title = "Accuracy and Rolling Standard Deviation of Accuracy per Window",
    x = "Window Number",
    y = "Value",
    color = "Metric"
  ) +
  scale_color_manual(values = c("Accuracy" = "navyblue", "Accuracy SD" = "orange", "Major" = "red")) +  # Custom colors including red for major
  theme_minimal()
```



## PREDICTING AGAINST THE BOOKS


```{r}
# merge the odds_data with the original predictive_dataframe
merged_original <- predictive_dataframe_pr %>%
  left_join(odds_data, by = c("EP_game_id", "player1", "player2")) %>%
  rename(
    golfer1_win_orig = golfer1_win,
    golfer2_win_orig = golfer2_win
  )

# swap player1 and player2 in odds_data for the merge
odds_data_swapped <- odds_data %>%
  rename(
    player1 = player2,
    player2 = player1,
    golfer1_win = golfer2_win,
    golfer2_win = golfer1_win
  )

# merge the odds_data with the predictive_dataframe with swapped order
merged_swapped <- predictive_dataframe_pr %>%
  left_join(odds_data_swapped, by = c("EP_game_id", "player1", "player2")) %>%
  rename(
    golfer1_win_swapped = golfer1_win,
    golfer2_win_swapped = golfer2_win
  )

# combine the results from the original and swapped merges
predictive_odds_data_pr <- merged_original %>%
  left_join(merged_swapped, by = c("EP_game_id", "player1", "player2", "player1_win",   "ranking_player1", "ranking_player2", "major"))

# coalesce the columns to handle NA values and get the correct golfer1_win and golfer2_win
predictive_odds_data_pr <- predictive_odds_data_pr %>%
  mutate(
    tournament = coalesce(tournament.x.x, tournament.x.y, tournament.y.x, tournament.y.y),
    golfer1_win = coalesce(golfer1_win_orig, golfer1_win_swapped),
    golfer2_win = coalesce(golfer2_win_orig, golfer2_win_swapped),
    prob_player1_win = coalesce(prob_player1_win.x, prob_player1_win.y),
    prob_player2_win = coalesce(prob_player2_win.x, prob_player2_win.y)
  ) %>%
  select(-golfer1_win_orig, -golfer2_win_orig, -golfer1_win_swapped, -golfer2_win_swapped, -tournament.x.x, -tournament.x.y, -tournament.y.x, -tournament.y.y, -prob_player1_win.x, -prob_player1_win.y, -prob_player2_win.x, -prob_player2_win.y, window_number.x, major) %>%
  rename(window_number = window_number.x)
```

Selecting relevant columns and deleting data used in the first window:

```{r}
predictive_odds_data_pr <- predictive_odds_data_pr %>%
  select(EP_game_id, tournament, player1, player2, golfer1_win, golfer2_win, prob_player1_win, prob_player2_win, player1_win, window_number, major)

predictive_odds_data_pr <- predictive_odds_data_pr %>%
  drop_na()
```

```{r}
dim(predictive_odds_data_pr)
```


Calculating accuracy of these predictions:

```{r}
correct_predictions <- predictive_odds_data_pr %>%
  mutate(predicted_win = ifelse(prob_player1_win > 0.5, 1, 0)) %>%
  summarize(accuracy = mean(predicted_win == player1_win))

print(correct_predictions)
```

Calculate log loss:

```{r}
log_loss <- predictive_odds_data_pr %>%
  mutate(log_loss = - (player1_win * log(prob_player1_win) + (1 - player1_win) * log(1 - prob_player1_win))) %>%
  summarize(mean_log_loss = mean(log_loss))

print(log_loss)
```
Calculating Brier Score:

```{r}
brier_score <- predictive_odds_data_pr %>%
  mutate(brier_score = (prob_player1_win - player1_win)^2) %>%
  summarize(mean_brier_score = mean(brier_score))

print(brier_score)
```

Adding a column with implied probabilities:

```{r}
predictive_odds_data_pr <- predictive_odds_data_pr %>%
  mutate(
    implied_prob_player1 = 1 / golfer1_win,
    implied_prob_player2 = 1 / golfer2_win
  )
```

Identify the value bets where model's predicted probability is greater than the bookmaker's implied probability:

```{r}
predictive_odds_data_pr <- predictive_odds_data_pr %>%
  mutate(
    value_bet_player1 = (prob_player1_win - implied_prob_player1) > 0,
    value_bet_player2 = (prob_player2_win - implied_prob_player2) > 0 
  )

# filtering out the dataframe for which both value_bets are FALSE:

predictive_odds_data_pr <- predictive_odds_data_pr %>%
  filter(value_bet_player1 == TRUE | value_bet_player2 == TRUE)
```

Adding Kelly Fractions:

```{r}
predictive_odds_data_pr <- predictive_odds_data_pr %>%
  mutate(
    b_player1 = golfer1_win -1,
    b_player2 = golfer2_win -1, 
         
    q_player1 = 1 - prob_player1_win,
    q_player2 = 1 - prob_player2_win,
    
    kelly_fraction_player1 = ifelse(value_bet_player1, 
                                 (b_player1 * prob_player1_win - q_player1) / b_player1, 
                                 0),
    kelly_fraction_player2 = ifelse(value_bet_player2, 
                                 (b_player2 * prob_player2_win - q_player2) / b_player2, 
                                 0)) %>%
  select(-b_player1, -b_player2, - q_player1, -q_player2)

```

Will it require a normalising constant for unit?

```{r}
matches_per_window <- predictive_odds_data_pr %>%
  group_by(window_number) %>%
  summarise(num_matches = n())
```

Implementing 3 Strategies:


```{r}
bankroll_kelly <- 1000
bankroll_half_kelly <- 1000
bankroll_unit <- 1000

# start with the initial bankroll, no stakes, no profit/loss, and no expected value
bankroll_kelly_progress <- c(bankroll_kelly)
stake_kelly_progress <- c(0)
profit_loss_kelly_progress <- c(0)
expected_value_kelly_progress <- c(0)

bankroll_half_kelly_progress <- c(bankroll_half_kelly)
stake_half_kelly_progress <- c(0)
profit_loss_half_kelly_progress <- c(0)
expected_value_half_kelly_progress <- c(0)

bankroll_unit_progress <- c(bankroll_unit)
stake_unit_progress <- c(0)
profit_loss_unit_progress <- c(0)
expected_value_unit_progress <- c(0)

# add an initial window 0 to represent the starting state
windows <- c(0, sort(unique(predictive_odds_data_pr$window_number)))

```

```{r}
for (window in windows[-1]) {  # skip the first window 0 as it's already initialized
    # filter bets for the current window
    bets_in_window <- subset(predictive_odds_data_pr, window_number == window)
    
    # calculate total Kelly fractions for normalisation if necessary
    total_fraction <- sum(bets_in_window$kelly_fraction_player1) + sum(bets_in_window$kelly_fraction_player2)
    if (total_fraction > 1) {
        normalization_factor <- 1 / total_fraction
        bets_in_window$kelly_fraction_player1 <- bets_in_window$kelly_fraction_player1 * normalization_factor
        bets_in_window$kelly_fraction_player2 <- bets_in_window$kelly_fraction_player2 * normalization_factor
    } else {
        normalization_factor <- 1
    }

    # reset stakes, profits/losses, and expected value for the current window
    stake_kelly_window <- 0
    profit_loss_kelly_window <- 0
    expected_value_kelly_window <- 0
    
    stake_half_kelly_window <- 0
    profit_loss_half_kelly_window <- 0
    expected_value_half_kelly_window <- 0
    
    stake_unit_window <- 0
    profit_loss_unit_window <- 0
    expected_value_unit_window <- 0
    
    # kelly Stake
    for (i in 1:nrow(bets_in_window)) {
        bet <- bets_in_window[i, ]
        
        if (bet$value_bet_player1) {
            stake_kelly <- bankroll_kelly * bet$kelly_fraction_player1
            stake_kelly_window <- stake_kelly_window + stake_kelly
            
            expected_value_kelly <- stake_kelly * bet$golfer1_win - stake_kelly * (1 - bet$golfer1_win)
            expected_value_kelly_window <- expected_value_kelly_window + expected_value_kelly
            
            if (bet$player1_win == 1) {
                profit_loss_kelly_window <- profit_loss_kelly_window + stake_kelly * (bet$golfer1_win - 1)
            } else {
                profit_loss_kelly_window <- profit_loss_kelly_window - stake_kelly
            }
        }
        
        if (bet$value_bet_player2) {
            stake_kelly <- bankroll_kelly * bet$kelly_fraction_player2
            stake_kelly_window <- stake_kelly_window + stake_kelly
            
            expected_value_kelly <- stake_kelly * bet$golfer2_win - stake_kelly * (1 - bet$golfer2_win)
            expected_value_kelly_window <- expected_value_kelly_window + expected_value_kelly
            
            if (bet$player1_win == 0) {
                profit_loss_kelly_window <- profit_loss_kelly_window + stake_kelly * (bet$golfer2_win - 1)
            } else {
                profit_loss_kelly_window <- profit_loss_kelly_window - stake_kelly
            }
        }
    }
    
    # update bankroll after the window
    bankroll_kelly <- bankroll_kelly + profit_loss_kelly_window

    # store the results for Kelly Stake
    bankroll_kelly_progress <- c(bankroll_kelly_progress, bankroll_kelly)
    stake_kelly_progress <- c(stake_kelly_progress, stake_kelly_window)
    profit_loss_kelly_progress <- c(profit_loss_kelly_progress, profit_loss_kelly_window)
    expected_value_kelly_progress <- c(expected_value_kelly_progress, expected_value_kelly_window)
    
    # half Kelly Stake
    for (i in 1:nrow(bets_in_window)) {
        bet <- bets_in_window[i, ]
        
        if (bet$value_bet_player1) {
            stake_half_kelly <- bankroll_half_kelly * (bet$kelly_fraction_player1 / 2)
            stake_half_kelly_window <- stake_half_kelly_window + stake_half_kelly
            
            expected_value_half_kelly <- stake_half_kelly * bet$golfer1_win - stake_half_kelly * (1 - bet$golfer1_win)
            expected_value_half_kelly_window <- expected_value_half_kelly_window + expected_value_half_kelly
            
            if (bet$player1_win == 1) {
                profit_loss_half_kelly_window <- profit_loss_half_kelly_window + stake_half_kelly * (bet$golfer1_win - 1)
            } else {
                profit_loss_half_kelly_window <- profit_loss_half_kelly_window - stake_half_kelly
            }
        }
        
        if (bet$value_bet_player2) {
            stake_half_kelly <- bankroll_half_kelly * (bet$kelly_fraction_player2 / 2)
            stake_half_kelly_window <- stake_half_kelly_window + stake_half_kelly
            
            expected_value_half_kelly <- stake_half_kelly * bet$golfer2_win - stake_half_kelly * (1 - bet$golfer2_win)
            expected_value_half_kelly_window <- expected_value_half_kelly_window + expected_value_half_kelly
            
            if (bet$player1_win == 0) {
                profit_loss_half_kelly_window <- profit_loss_half_kelly_window + stake_half_kelly * (bet$golfer2_win - 1)
            } else {
                profit_loss_half_kelly_window <- profit_loss_half_kelly_window - stake_half_kelly
            }
        }
    }
    
    bankroll_half_kelly <- bankroll_half_kelly + profit_loss_half_kelly_window
    
    bankroll_half_kelly_progress <- c(bankroll_half_kelly_progress, bankroll_half_kelly)
    stake_half_kelly_progress <- c(stake_half_kelly_progress, stake_half_kelly_window)
    profit_loss_half_kelly_progress <- c(profit_loss_half_kelly_progress, profit_loss_half_kelly_window)
    expected_value_half_kelly_progress <- c(expected_value_half_kelly_progress, expected_value_half_kelly_window)
    
    # unit Stake
    total_unit_fraction <- nrow(bets_in_window) * 0.05
    if (total_unit_fraction > 1) {
        normalization_factor <- 1 / total_unit_fraction
    } else {
        normalization_factor <- 1
    }
    
    for (i in 1:nrow(bets_in_window)) {
        bet <- bets_in_window[i, ]
        stake_unit <- bankroll_unit * 0.05 * normalization_factor
        stake_unit_window <- stake_unit_window + stake_unit
        
        expected_value_unit <- stake_unit * bet$golfer1_win - stake_unit * (1 - bet$golfer1_win)
        expected_value_unit_window <- expected_value_unit_window + expected_value_unit
        
        if (bet$value_bet_player1) {
            if (bet$player1_win == 1) {
                profit_loss_unit_window <- profit_loss_unit_window + stake_unit * (bet$golfer1_win - 1)
            } else {
                profit_loss_unit_window <- profit_loss_unit_window - stake_unit
            }
        }
        
        if (bet$value_bet_player2) {
            if (bet$player1_win == 0) {
                profit_loss_unit_window <- profit_loss_unit_window + stake_unit * (bet$golfer2_win - 1)
            } else {
                profit_loss_unit_window <- profit_loss_unit_window - stake_unit
            }
        }
    }
    
    bankroll_unit <- bankroll_unit + profit_loss_unit_window
    
    bankroll_unit_progress <- c(bankroll_unit_progress, bankroll_unit)
    stake_unit_progress <- c(stake_unit_progress, stake_unit_window)
    profit_loss_unit_progress <- c(profit_loss_unit_progress, profit_loss_unit_window)
    expected_value_unit_progress <- c(expected_value_unit_progress, expected_value_unit_window)
}
```

Datframe for all the important metrics:

```{r}
results_pr <- data.frame(
    window = windows,
    bankroll_kelly = bankroll_kelly_progress,
    bankroll_half_kelly = bankroll_half_kelly_progress,
    bankroll_unit = bankroll_unit_progress,
    stake_kelly = stake_kelly_progress,
    stake_half_kelly = stake_half_kelly_progress,
    stake_unit = stake_unit_progress,
    profit_loss_kelly = profit_loss_kelly_progress,
    profit_loss_half_kelly = profit_loss_half_kelly_progress,
    profit_loss_unit = profit_loss_unit_progress,
    expected_value_kelly = expected_value_kelly_progress,
    expected_value_half_kelly = expected_value_half_kelly_progress,
    expected_value_unit = expected_value_unit_progress
)

# adding majors 
major_windows <- c(14, 21, 27, 35)
results_pr$major <- ifelse(results_pr$window %in% major_windows, 1, 0)

majors_profit_loss <- results_pr %>%
  filter(major == 1) %>%  # filter only major tournaments
  select(profit_loss_kelly, profit_loss_half_kelly, profit_loss_unit, window)  # Select profit and loss columns

#calculate EV ROI
results_pr$ev_roi_kelly <- results_pr$expected_value_kelly / results_pr$stake_kelly * 100
results_pr$ev_roi_half_kelly <- results_pr$expected_value_half_kelly / results_pr$stake_half_kelly * 100
results_pr$ev_roi_unit <- results_pr$expected_value_unit / results_pr$stake_unit * 100

# calculate Actual ROI
results_pr$roi_kelly <- results_pr$profit_loss_kelly / results_pr$stake_kelly * 100
results_pr$roi_half_kelly <- results_pr$profit_loss_half_kelly / results_pr$stake_half_kelly * 100
results_pr$roi_unit <- results_pr$profit_loss_unit / results_pr$stake_unit * 100

results_pr[is.na(results_pr) | results_pr == Inf] <- 0

View(results_pr)
```

Bankroll progression: 

```{r}
ggplot(results_pr, aes(x = window)) +
    geom_line(aes(y = bankroll_kelly, color = "Kelly Stake")) +
    geom_line(aes(y = bankroll_half_kelly, color = "Half Kelly Stake")) +
    geom_line(aes(y = bankroll_unit, color = "Unit Stake")) +
    labs(
        title = "Rolling Bankroll Over Time",
        x = "Window",
        y = "Bankroll",
        color = "Strategy"
    ) +
    scale_color_manual(values = c("Kelly Stake" = "blue", "Half Kelly Stake" = "red", "Unit Stake" = "green")) +
    theme_minimal()
```

Actual ROI vs Expected ROI:

```{r}
filtered_results_pr <- results_pr %>% filter(window >= 1)

ggplot(filtered_results_pr, aes(x = window)) +
    geom_line(aes(y = roi_kelly, color = "Kelly Stake", linetype = "Actual")) +
    geom_line(aes(y = ev_roi_kelly, color = "Kelly Stake", linetype = "Expected")) +
    geom_line(aes(y = roi_unit, color = "Unit Stake", linetype = "Actual")) +
    geom_line(aes(y = ev_roi_unit, color = "Unit Stake", linetype = "Expected")) +
    scale_color_manual(values = c("Kelly Stake" = "blue", "Unit Stake" = "red")) +  # Custom colors
    labs(title = "Actual vs. Expected ROI Over Time",
         x = "Window",
         y = "ROI (%)",
         color = "Staking Strategy",
         linetype = "ROI Type") +
    theme_minimal()
```

Cumulative Profit / Loss:

```{r}
results_pr$cumulative_profit_loss_kelly <- cumsum(results_pr$profit_loss_kelly)
results_pr$cumulative_profit_loss_half_kelly <- cumsum(results_pr$profit_loss_half_kelly)
results_pr$cumulative_profit_loss_unit <- cumsum(results_pr$profit_loss_unit)

ggplot(results_pr, aes(x = window)) +
    geom_line(aes(y = cumulative_profit_loss_kelly, color = "Kelly Stake")) +
    geom_line(aes(y = cumulative_profit_loss_half_kelly, color = "Half Kelly Stake")) +
    geom_line(aes(y = cumulative_profit_loss_unit, color = "Unit Stake")) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "steelblue") +
    labs(title = "Cumulative Profit/Loss Over Time",
         x = "Window",
         y = "Cumulative Profit/Loss",
         color = "Staking Strategy") +
  scale_color_manual(values = c("Kelly Stake" = "blue", "Half Kelly Stake" = "green", "Unit Stake" = "red")) +
    theme_minimal()
```


Boxplot of ROI across:

```{r}
results_melted <- reshape2::melt(results_pr, id.vars = "window", measure.vars = c("roi_kelly", "roi_unit"))

ggplot(results_melted, aes(x = variable, y = value, fill = variable)) +
    geom_boxplot() +
    labs(title = "Distribution of ROI for Each Strategy",
         x = "Staking Strategy",
         y = "ROI (%)",
         fill = "Staking Strategy") +
    theme_minimal()
```





