---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/lukeconkleton/Library/Mobile Documents/com~apple~CloudDocs/Documents/UCL Project/Data")
```

```{r, include=FALSE}
library(tidyverse)
library(BradleyTerry2)
library(tidyr)
library(zoo)
```

```{r}
rounds_data <- readRDS("roundData2018-2022.rds")
odds_data <- readRDS("oddsData2018-2022.rds")
```

## 1. DATA WRANGLING & FEATURE SELECTION

Subsetting dataframe to only include pre-covid data. Creating a new dataframe to only inlcude variables of interest; golfer, tournament, major, final position and date.

```{r}
rounds_data <- rounds_data %>% filter(year %in% c('2018', '2019'))
```

Engineering a feature for whether or not the golfer is playing at 'home':

```{r}
rounds_data$is.home <- ifelse(rounds_data$player_country == rounds_data$tournament_country, 1, 0)
```

```{r}
rounds <- data.frame(
  EP_game_id = rounds_data$EP_game_id,
  tournament = rounds_data$event_name,
  major = rounds_data$is_major,
  golfer_id = rounds_data$player_id,
  golfer = rounds_data$player_name,
  final_position = rounds_data$final_position,
  start_date = rounds_data$competition_start_date, 
  times_played_course = rounds_data$times_played_course,
  is.home = rounds_data$is.home
)

rounds <- rounds %>%
  distinct()
```

Only keeping a variable for times played course pre-tournament.

```{r}
rounds <- rounds %>%
  group_by(
    EP_game_id, tournament, golfer_id, start_date) %>%
  slice_min(
    times_played_course) %>% #takes the minimum value of the datapoints
  ungroup() %>%
  rename(
    times_played_course_pre_tournament = times_played_course)
```

Only considering the relevant time frame:

```{r}
odds_data$competition_start_date <- as.Date(odds_data$competition_start_date)

odds_data <- odds_data %>% filter(format(competition_start_date, "%Y") %in% c("2018", "2019"))

odds_data <- odds_data %>%
  select(EP_game_id, event_name, SMARTODDS_golfer1_name, SMARTODDS_golfer2_name, golfer1_win, golfer2_win) %>%
  rename(
    player1 = SMARTODDS_golfer1_name,
    player2 = SMARTODDS_golfer2_name,
    tournament = event_name
  )
```

More tournaments included in rounds data set comparatively:

```{r}
length(unique(odds_data$EP_game_id))
length(unique(rounds_data$EP_game_id))
```

How many times does each golfer appear in the odds dataframe?

Subsetting the rounds dataset to only include golfers who have made over 50 appearances in the odds dataset:

```{r}
odds_golfers <- c(odds_data$player1, odds_data$player2)
length(unique(odds_golfers))

# how many appearances in the odds data frame
golfer_counts <- data.frame(golfer = odds_golfers) %>%
  group_by(golfer) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

players_over_50 <- golfer_counts %>%
  filter(count >= 50)

dim(players_over_50)
```


```{r}
# filter the original dataframe to keep only those players
filtered_rounds_data <- rounds_data %>%
  filter(player_name %in% players_over_50$golfer)

# filter Rounds Dataframe

filtered_rounds <- rounds %>%
  filter(golfer %in% players_over_50$golfer) %>%
  filter(!is.na(final_position) & final_position != 0)
```

## 2. CREATING A PAIRWISE COMPARISON DATASET

Create a dataset with every single match-up in that time period. Can be modified to be grouped via 'EP_game_ID' or by 'tournament'. Similarly, is it easier to use indexes for the golfers or their names? For tournaments, need to sort by ID as there will be repetitions (two years means 2x each tournament).


```{r}
# define the create_pairs function
create_pairs <- function(df) {
  pairs <- expand.grid(golfer_A = df$golfer_id, golfer_B = df$golfer_id)
  pairs <- pairs[pairs$golfer_A != pairs$golfer_B, ]
  
  golfer_A_indices <- match(pairs$golfer_A, df$golfer_id)
  golfer_B_indices <- match(pairs$golfer_B, df$golfer_id)
  
  if (length(golfer_A_indices) == 0 || length(golfer_B_indices) == 0) {
    return(data.frame()) 
  }
  
  result <- data.frame(
    golfer_A = df$golfer[golfer_A_indices],
    golfer_B = df$golfer[golfer_B_indices],
    final_position_A = df$final_position[golfer_A_indices],
    final_position_B = df$final_position[golfer_B_indices],
    start_date = df$start_date[golfer_A_indices],
    major = df$major[golfer_A_indices],
    EP_game_id = df$EP_game_id[golfer_A_indices],
    tournament = df$tournament[golfer_A_indices],
    playerA_times_played = df$times_played_course_pre_tournament[golfer_A_indices],
    playerB_times_played = df$times_played_course_pre_tournament[golfer_B_indices],
    is.home.A = df$is.home[golfer_A_indices],
    is.home.B = df$is.home[golfer_B_indices]
  )
  
  return(result)
}

# apply the function to each tournament
pairwise_matches <- filtered_rounds %>%
  group_by(EP_game_id) %>%
  do({
    result <- create_pairs(.)
    if (nrow(result) > 0) {
      result$EP_game_id <- .$EP_game_id[1]
    }
    result
  }) %>%
  ungroup()

pairwise_matches <- pairwise_matches %>%
  mutate(
    result_A = ifelse(final_position_A < final_position_B, 1, 0),
    result_B = ifelse(final_position_B < final_position_A, 1, 0)
  )

```

Ensure that there are no duplicate results in the dataset and that the data is ordered temporally. Will need to be amended if adding variables.

```{r}
# ensure each pair is in a consistent order
pairwise_matches <- pairwise_matches %>%
  rowwise() %>%
  mutate(
    golfer_A_sorted = ifelse(golfer_A < golfer_B, golfer_A, golfer_B),
    golfer_B_sorted = ifelse(golfer_A < golfer_B, golfer_B, golfer_A),
    final_position_A_sorted = ifelse(golfer_A < golfer_B, final_position_A, final_position_B),
    final_position_B_sorted = ifelse(golfer_A < golfer_B, final_position_B, final_position_A),
    result_A_sorted = ifelse(golfer_A < golfer_B, result_A, result_B),
    result_B_sorted = ifelse(golfer_A < golfer_B, result_B, result_A),
    is.home_A_sorted = ifelse(golfer_A < golfer_B, is.home.A, is.home.B),
    is.home_B_sorted = ifelse(golfer_A < golfer_B, is.home.B, is.home.A),
    times_played_A_sorted = ifelse(golfer_A < golfer_B, playerA_times_played, playerB_times_played),
    times_played_B_sorted = ifelse(golfer_A < golfer_B, playerB_times_played, playerA_times_played)
  ) %>%
  ungroup()

# add home advantage columns after sorting
pairwise_matches <- pairwise_matches %>%
  mutate(
    home_advantage_A = ifelse(is.home_A_sorted == 1 & is.home_B_sorted == 0, 1, 0),
    home_advantage_B = ifelse(is.home_A_sorted == 0 & is.home_B_sorted == 1, 1, 0)
  )

# remove duplicates based on the sorted pairs and the tournament
pairwise_matches <- pairwise_matches %>%
  select(EP_game_id, tournament, golfer_A_sorted, golfer_B_sorted, final_position_A_sorted, final_position_B_sorted, 
         result_A_sorted, result_B_sorted, start_date, major, times_played_A_sorted, times_played_B_sorted, 
         is.home_A_sorted, is.home_B_sorted, home_advantage_A, home_advantage_B) %>%
  distinct() %>%
  rename(
    golfer_A = golfer_A_sorted,
    golfer_B = golfer_B_sorted,
    final_position_A = final_position_A_sorted,
    final_position_B = final_position_B_sorted,
    result_A = result_A_sorted,
    result_B = result_B_sorted,
    is.home_A = is.home_A_sorted,
    is.home_B = is.home_B_sorted,
    times_played_A = times_played_A_sorted,
    times_played_B = times_played_B_sorted
  )

# ensuring proper column naming for clarity
pairwise_matches <- pairwise_matches %>%
  rename(
    player1 = golfer_A,
    player2 = golfer_B,
    score1 = result_A,
    score2 = result_B
  )

# maintaining temporal ordering
pairwise_matches <- pairwise_matches[order(pairwise_matches$start_date), ]
```

## 3. CREATING A ROLLING WINDOW

Utilising a rolling window; takes a subset of the data and moves the window across the dataset with the window size constant(ish). The window slides forward by an observation at each step.

In this, an 'observation' is a tournamnet, so the window won't be necessarily have the same number of datapoints but will have the same number of tournaments.

This allows for the dynamic updating of the model as new data becomes available.

```{r}
unique_tournaments <- unique(pairwise_matches$EP_game_id)
length(unique_tournaments)
```

This function seems to be working.

```{r}
create_rolling_windows <- function(data, tournaments, window_size) {
  n <- length(tournaments)
  windows <- list()
  
  for (start in 1:(n - window_size + 1)) {
    end <- start + window_size - 1
    window_tournaments <- tournaments[start:end]
    window_data <- data[data$EP_game_id %in% window_tournaments, ]
    windows[[length(windows) + 1]] <- window_data
  }
  
  return(windows)
}

# window size (number of tournaments per window)
window_size <- 80

windows <- create_rolling_windows(pairwise_matches, unique_tournaments, window_size)
```

Create a window dataframe.

```{r}
create_window_summary <- function(windows) {
  summary_df <- data.frame(window_number = integer(),
                           final_EP_game_id = integer(),
                           final_tournament = character(),
                           start_data = character(),
                           major = integer(),
                           stringsAsFactors = FALSE)
  
  for (i in 1:(length(windows) - 1)) {
    next_window <- windows[[i + 1]]
    final_EP_game_id <- tail(next_window$EP_game_id, n = 1)  
    final_tournament <- tail(next_window$tournament, n = 1)  
    final_start_date <- tail(next_window$start_date, n = 1)  
    final_major <- tail(next_window$major, n = 1)  
    
    summary_df <- rbind(summary_df, data.frame(window_number = i,
                                               next_EP_game_id = final_EP_game_id,
                                               next_tournament = final_tournament,
                                               start_date = final_start_date,
                                               major = final_major))
  }
  
  return(summary_df)
}

window_summary <- create_window_summary(windows)
```

## 4. PREPARING DATA FOR TRAINING

Prepare the data so it is compatible with the Bradley-Terry package.

```{r}
prepare_pairwise_data <- function(data) {
  players <- unique(c(data$player1, data$player2))
  
  pairwise_data <- data.frame(
    player1 = factor(data$player1, levels = players),
    player2 = factor(data$player2, levels = players),
    wins1 = data$score1,
    wins2 = data$score2,
    major = data$major,
    played_before1 = data$times_played_A,
    played_before2 = data$times_played_B,
    is.home1 = data$is.home_A,
    is.home2 = data$is.home_B
  )
  
  pairwise_data$player1 <- data.frame(
  player = pairwise_data$player1,
  times_played = pairwise_data$played_before1,
  is.home = pairwise_data$is.home1
)

pairwise_data$player2 <- data.frame(
  player = pairwise_data$player2,
  times_played = pairwise_data$played_before2,
  is.home = pairwise_data$is.home2
)
  
  return(pairwise_data)
}

```


## 5. TRAINING MODEL

```{r}
train_bradley_terry2 <- function(window_data, window_number, ref_player) {
  pairwise_data <- prepare_pairwise_data(window_data)
  cat("Training model for window", window_number, "\n")
  model <- BTm(cbind(wins1, wins2),
               formula = ~ player + is.home,
               player1 = player1, 
               player2 = player2, 
               data = pairwise_data,
               id = "player",
               refcat = ref_player)
  return(model)
}

reference_player <- "Tommy Fleetwood" # basis of comparison

models2 <- lapply(seq_len(min(100, length(windows))), function(i) {
  train_bradley_terry2(windows[[i]], i, reference_player)
})

```


Creating a dataframe comparing the significance levels through all the windows:

```{r}
extract_significance <- function(model_summary) {

  coefs <- summary(model_summary)$coefficients
  
  # extract significance values for 'is.home'
  is_home_p_value <- coefs["is.home", "Pr(>|z|)"]
  
  return(is_home = is_home_p_value)
}
```

Extracting from relevant model:

```{r}
significance_values_2 <- lapply(models2, function(model) {
  extract_significance(model)
})

significance_df_2 <- do.call(rbind, significance_values_2)
colnames(significance_df_2) <- "is_home_p_value"

significance_df_2 <- as.data.frame(significance_df_2)
```

Checking mean of p-values:

```{r}
mean(significance_df_2$is_home_p_value, na.rm = TRUE)
```

Checking percentage of p-values under 5:

```{r}
percentage_under_threshold <- function(column, threshold = 0.05) {
  mean(column < threshold) * 100
}

percentage_signif <- significance_df_2 %>%
  summarize(across(everything(), percentage_under_threshold))

print(percentage_signif)
```

## Creating dataframe for coefficient values

```{r}
# extract coefficients from each model and store them in a list of DataFrames
extract_coefficients <- function(models, window_numbers, ref_player) {
  coef_list <- lapply(window_numbers, function(i) {
    if (i <= length(models)) {
      model <- models[[i]]
      coefs <- coef(model)
      if (!(ref_player %in% names(coefs))) {
        coefs <- c(coefs, setNames(0, ref_player))
      }
      data.frame(
        Player = names(coefs),
        Coefficient = as.numeric(coefs),
        Window = paste("Window", i, sep = "_"),
        stringsAsFactors = FALSE
      )
    } else {
      data.frame(
        Player = character(0),
        Coefficient = numeric(0),
        Window = character(0),
        stringsAsFactors = FALSE
      )
    }
  })
  return(coef_list)
}

```

Running the function for this model:

```{r}
window_numbers <- seq_len(length(models2))
coef_list <- extract_coefficients(models2, window_numbers, reference_player)

# combine the list of DataFrames into a single DataFrame
combined_df <- bind_rows(coef_list)

# ensure the reference player is in the final DataFrame
if (!("Tommy Fleetwood" %in% combined_df$Player)) {
  for (i in window_numbers) {
    combined_df <- combined_df %>%
      add_row(Player = reference_player, Coefficient = 0, Window = paste("Window", i, sep = "_"))
  }
}

final_df_2 <- combined_df %>%
  pivot_wider(names_from = Window, values_from = Coefficient) %>%
  arrange(Player)

# remove leading dots from the 'Player' column in final_df
final_df_2$Player <- gsub("^player", "", final_df_2$Player)

View(final_df_2)
```

## PREDICTIVE ANALYSIS

```{r}
predictive_dataframe2 <- merge(pairwise_matches, window_summary, 
                       by.x = "EP_game_id", by.y = "next_EP_game_id", all.x = TRUE) %>%
  select(window_number, start_date.x, EP_game_id, tournament, player1, player2, final_position_A, final_position_B, score1, score2, times_played_A, times_played_B, is.home_A, is.home_B, major.x) %>%
  na.omit() %>% # gets rid of all data prior to window 1
  rename(major = major.x, start_date = start_date.x, player1_win = score1)
```

Now merging strengths and covariate coefficient values:

```{r}
# merging the strengths 
predictive_dataframe2 <- predictive_dataframe2 %>%
  left_join(final_df_2 %>% 
              gather(Window, Strength, -Player) %>% 
              mutate(Window = as.numeric(gsub("Window_", "", Window))), 
            by = c("window_number" = "Window", "player1" = "Player")) %>%
  rename(player1_strength = Strength) %>%
  left_join(final_df_2 %>% 
              gather(Window, Strength, -Player) %>% 
              mutate(Window = as.numeric(gsub("Window_", "", Window))), 
            by = c("window_number" = "Window", "player2" = "Player")) %>%
  rename(player2_strength = Strength)

```

```{r}
# identify the indices or rows for the covariates
is_home_row <- which(final_df_2$Player == "is.home")

# extracting the data for merging
is_home_coeffs <- final_df_2[is_home_row, ] %>%
  gather(Window, is_home_coeff, -Player) %>%
  mutate(Window = as.numeric(gsub("Window_", "", Window)))

# merging the coefficients into the predictive dataframe
predictive_dataframe2 <- predictive_dataframe2 %>%
  left_join(is_home_coeffs, by = c("window_number" = "Window"))
```

Deselecting certain boys:

```{r}
predictive_dataframe2 <- predictive_dataframe2 %>%
  select(-start_date, -final_position_A, -final_position_B, -Player)
```

Adding columns for model probabilities:

```{r}
logistic <- function(x){
  1 / (1+exp(-x))
}

predictive_dataframe2 <- predictive_dataframe2 %>%
    mutate(
    linear_predictor_1 = player1_strength - player2_strength + 
                         is_home_coeff * (is.home_A - is.home_B),
    linear_predictor_2 = player2_strength - player1_strength + 
                         is_home_coeff * (is.home_B - is.home_A),
    prob_player1_win = logistic(linear_predictor_1),
    prob_player2_win = logistic(linear_predictor_2)
  ) %>%
  select(-linear_predictor_1, -linear_predictor_2)
```

```{r}
# get rid of all non-predictive data
predictive_dataframe2 <- predictive_dataframe2 %>%
  drop_na()

correct_predictions <- predictive_dataframe2 %>%
  mutate(predicted_win = ifelse(prob_player1_win > 0.5, 1, 0)) %>%
  summarize(accuracy = mean(predicted_win == player1_win))

print(correct_predictions)
```
Calculating log loss:

```{r}
log_loss <- predictive_dataframe2 %>%
  mutate(log_loss = - (player1_win * log(prob_player1_win) + (1 - player1_win) * log(1 - prob_player1_win))) %>%
  summarize(mean_log_loss = mean(log_loss))

print(log_loss)
```
Calculating Brier Score:

```{r}
brier_score <- predictive_dataframe2 %>%
  mutate(brier_score = (prob_player1_win - player1_win)^2) %>%
  summarize(mean_brier_score = mean(brier_score))

print(brier_score)
```
Accuracy per window:

```{r}
predictive_dataframe2 <- predictive_dataframe2 %>%
    mutate(predicted_win = ifelse(prob_player1_win > 0.5, 1, 0))

accuracy_per_window_2 <- predictive_dataframe2 %>%
    group_by(window_number) %>%  # Group by the window number
    summarize(accuracy = mean(predicted_win == player1_win))

print(accuracy_per_window_2)

ggplot(accuracy_per_window_2, aes(x = window_number, y = accuracy)) +
    geom_line(color = "blue") +
    geom_point(color = "red") +
    labs(
        title = "Accuracy per Window",
        x = "Window Number",
        y = "Accuracy"
    ) +
    theme_minimal()
```

##8. PROFITABILITY ANALYSIS

Merging the odds dataframe with the predictive dataframe:

```{r}
# merge the odds_data with the original predictive_dataframe
merged_original <- predictive_dataframe2 %>%
  left_join(odds_data, by = c("EP_game_id", "player1", "player2")) %>%
  rename(
    golfer1_win_orig = golfer1_win,
    golfer2_win_orig = golfer2_win
  )

# swap player1 and player2 in odds_data for the merge
odds_data_swapped <- odds_data %>%
  rename(
    player1 = player2,
    player2 = player1,
    golfer1_win = golfer2_win,
    golfer2_win = golfer1_win
  )

# merge the odds_data with the predictive_dataframe with swapped order
merged_swapped <- predictive_dataframe2 %>%
  left_join(odds_data_swapped, by = c("EP_game_id", "player1", "player2")) %>%
  rename(
    golfer1_win_swapped = golfer1_win,
    golfer2_win_swapped = golfer2_win
  )

# combine the results from the original and swapped merges
predictive_odds_data2 <- merged_original %>%
  left_join(merged_swapped, by = c("EP_game_id", "player1", "player2", "player1_win",   "player1_strength", "player2_strength"))

# coalesce the columns to handle NA values and get the correct golfer1_win and golfer2_win
predictive_odds_data2 <- predictive_odds_data2 %>%
  mutate(
    tournament = coalesce(tournament.x.x, tournament.x.y, tournament.y.x, tournament.y.y),
    golfer1_win = coalesce(golfer1_win_orig, golfer1_win_swapped),
    golfer2_win = coalesce(golfer2_win_orig, golfer2_win_swapped),
    prob_player1_win = coalesce(prob_player1_win.x, prob_player1_win.y),
    prob_player2_win = coalesce(prob_player2_win.x, prob_player2_win.y)
  ) %>%
  select(-golfer1_win_orig, -golfer2_win_orig, -golfer1_win_swapped, -golfer2_win_swapped, -tournament.x.x, -tournament.x.y, -tournament.y.x, -tournament.y.y, -prob_player1_win.x, -prob_player1_win.y, -prob_player2_win.x, -prob_player2_win.y)
```

Selecting relevant columns and deleting data used in the first window:

```{r}
predictive_odds_data2 <- predictive_odds_data2 %>%
  select(EP_game_id, tournament, player1, player2, golfer1_win, golfer2_win, prob_player1_win, prob_player2_win, player1_win )

predictive_odds_data2 <- predictive_odds_data2 %>%
  drop_na()
```

Calculating dimensions:

```{r}
dim(predictive_odds_data2)
```

Calculating Accuracy:

```{r}
correct_predictions <- predictive_odds_data2 %>%
  mutate(predicted_win = ifelse(prob_player1_win > 0.5, 1, 0)) %>%
  summarize(accuracy = mean(predicted_win == player1_win))

print(correct_predictions)
```
Calculate log loss:

```{r}
log_loss <- predictive_odds_data2 %>%
  mutate(log_loss = - (player1_win * log(prob_player1_win) + (1 - player1_win) * log(1 - prob_player1_win))) %>%
  summarize(mean_log_loss = mean(log_loss))

print(log_loss)
```
Calculating Brier Score:

```{r}
brier_score <- predictive_odds_data2 %>%
  mutate(brier_score = (prob_player1_win - player1_win)^2) %>%
  summarize(mean_brier_score = mean(brier_score))

print(brier_score)
```



