---
title: "PageRank (time-decay)"
output: html_document
date: "2024-08-13"
---


```{r, include=FALSE}
library(tidyverse)
library(igraph)
```

#############################################################################
#############################################################################

```{r}
rounds_data <- readRDS("roundData2018-2022.rds")
odds_data <- readRDS("oddsData2018-2022.rds")
```

## 1. DATA WRANGLING & FEATURE SELECTION

Subsetting dataframe to only include pre-covid data. Creating a new dataframe to only inlcude variables of interest; golfer, tournament, major, final position and date.

```{r}
rounds_data <- rounds_data %>% filter(year %in% c('2018', '2019'))
```

Engineering a feature for whether or not the golfer is playing at 'home':

```{r}
rounds_data$is.home <- ifelse(rounds_data$player_country == rounds_data$tournament_country, 1, 0)
```

```{r}
rounds <- data.frame(
  EP_game_id = rounds_data$EP_game_id,
  tournament = rounds_data$event_name,
  major = rounds_data$is_major,
  golfer_id = rounds_data$player_id,
  golfer = rounds_data$player_name,
  final_position = rounds_data$final_position,
  start_date = rounds_data$competition_start_date, 
  times_played_course = rounds_data$times_played_course,
  is.home = rounds_data$is.home
)

rounds <- rounds %>%
  distinct()
```

Only keeping a variable for times played course pre-tournament.

```{r}
rounds <- rounds %>%
  group_by(EP_game_id, tournament, golfer_id, start_date) %>%
  slice_min(times_played_course) %>% #takes the minimum value of the datapoints
  ungroup() %>%
  rename(times_played_course_pre_tournament = times_played_course)
```



Only considering the relevant time frame:

```{r}
odds_data$competition_start_date <- as.Date(odds_data$competition_start_date)

odds_data <- odds_data %>% filter(format(competition_start_date, "%Y") %in% c("2018", "2019"))

odds_data <- odds_data %>%
  select(EP_game_id, event_name, SMARTODDS_golfer1_name, SMARTODDS_golfer2_name, golfer1_win, golfer2_win) %>%
  rename(
    player1 = SMARTODDS_golfer1_name,
    player2 = SMARTODDS_golfer2_name,
    tournament = event_name
  )
```

More tournaments included in rounds data set comparatively:

```{r}
length(unique(odds_data$EP_game_id))
length(unique(rounds_data$EP_game_id))
```

How many times does each golfer appear in the odds dataframe?

a)  Subsetting the rounds dataset to only include golfers who have made over 'x' appearances in the odds dataset:

```{r}
odds_golfers <- c(odds_data$player1, odds_data$player2)
length(unique(odds_golfers))

# how many appearances in the odds data frame
golfer_counts <- data.frame(golfer = odds_golfers) %>%
  group_by(golfer) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

players_over_50 <- golfer_counts %>%
  filter(count >= 50)

dim(players_over_50)
```


```{r}
# filter the original dataframe to keep only those players
filtered_rounds_data <- rounds_data %>%
  filter(player_name %in% players_over_50$golfer)

# filter Rounds Dataframe

filtered_rounds <- rounds %>%
  filter(golfer %in% players_over_50$golfer) %>%
  filter(!is.na(final_position) & final_position != 0)
```

## 2. CREATING A PAIRWISE COMPARISON DATASET

Create a dataset with every single match-up in that time period.


```{r}
# define the create_pairs function
create_pairs <- function(df) {
  pairs <- expand.grid(golfer_A = df$golfer_id, golfer_B = df$golfer_id)
  pairs <- pairs[pairs$golfer_A != pairs$golfer_B, ]
  
  golfer_A_indices <- match(pairs$golfer_A, df$golfer_id)
  golfer_B_indices <- match(pairs$golfer_B, df$golfer_id)
  
  if (length(golfer_A_indices) == 0 || length(golfer_B_indices) == 0) {
    return(data.frame()) 
  }
  
  result <- data.frame(
    golfer_A = df$golfer[golfer_A_indices],
    golfer_B = df$golfer[golfer_B_indices],
    final_position_A = df$final_position[golfer_A_indices],
    final_position_B = df$final_position[golfer_B_indices],
    start_date = df$start_date[golfer_A_indices],
    major = df$major[golfer_A_indices],
    EP_game_id = df$EP_game_id[golfer_A_indices],
    tournament = df$tournament[golfer_A_indices],
    playerA_times_played = df$times_played_course_pre_tournament[golfer_A_indices],
    playerB_times_played = df$times_played_course_pre_tournament[golfer_B_indices],
    is.home.A = df$is.home[golfer_A_indices],
    is.home.B = df$is.home[golfer_B_indices]
  )
  
  return(result)
}

# apply the function to each tournament
pairwise_matches <- filtered_rounds %>%
  group_by(EP_game_id) %>%
  do({
    result <- create_pairs(.)
    if (nrow(result) > 0) {
      result$EP_game_id <- .$EP_game_id[1]
    }
    result
  }) %>%
  ungroup()

pairwise_matches <- pairwise_matches %>%
  mutate(
    result_A = ifelse(final_position_A < final_position_B, 1, 0),
    result_B = ifelse(final_position_B < final_position_A, 1, 0)
  )

```

Ensure that there are no duplicate results in the dataset and that the data is ordered temporally. Will need to be amended if adding variables.

```{r}
# ensure each pair is in a consistent order
pairwise_matches <- pairwise_matches %>%
  rowwise() %>%
  mutate(
    golfer_A_sorted = ifelse(golfer_A < golfer_B, golfer_A, golfer_B),
    golfer_B_sorted = ifelse(golfer_A < golfer_B, golfer_B, golfer_A),
    final_position_A_sorted = ifelse(golfer_A < golfer_B, final_position_A, final_position_B),
    final_position_B_sorted = ifelse(golfer_A < golfer_B, final_position_B, final_position_A),
    result_A_sorted = ifelse(golfer_A < golfer_B, result_A, result_B),
    result_B_sorted = ifelse(golfer_A < golfer_B, result_B, result_A),
    is.home_A_sorted = ifelse(golfer_A < golfer_B, is.home.A, is.home.B),
    is.home_B_sorted = ifelse(golfer_A < golfer_B, is.home.B, is.home.A),
    times_played_A_sorted = ifelse(golfer_A < golfer_B, playerA_times_played, playerB_times_played),
    times_played_B_sorted = ifelse(golfer_A < golfer_B, playerB_times_played, playerA_times_played)
  ) %>%
  ungroup()

# add home advantage columns after sorting
pairwise_matches <- pairwise_matches %>%
  mutate(
    home_advantage_A = ifelse(is.home_A_sorted == 1 & is.home_B_sorted == 0, 1, 0),
    home_advantage_B = ifelse(is.home_A_sorted == 0 & is.home_B_sorted == 1, 1, 0)
  )

# remove duplicates based on the sorted pairs and the tournament
pairwise_matches <- pairwise_matches %>%
  select(EP_game_id, tournament, golfer_A_sorted, golfer_B_sorted, final_position_A_sorted, final_position_B_sorted, 
         result_A_sorted, result_B_sorted, start_date, major, times_played_A_sorted, times_played_B_sorted, 
         is.home_A_sorted, is.home_B_sorted, home_advantage_A, home_advantage_B) %>%
  distinct() %>%
  rename(
    golfer_A = golfer_A_sorted,
    golfer_B = golfer_B_sorted,
    final_position_A = final_position_A_sorted,
    final_position_B = final_position_B_sorted,
    result_A = result_A_sorted,
    result_B = result_B_sorted,
    is.home_A = is.home_A_sorted,
    is.home_B = is.home_B_sorted,
    times_played_A = times_played_A_sorted,
    times_played_B = times_played_B_sorted
  )

pairwise_matches <- pairwise_matches %>%
  rename(
    player1 = golfer_A,
    player2 = golfer_B,
    score1 = result_A,
    score2 = result_B
  )

# maintaining temporal ordering
pairwise_matches <- pairwise_matches %>%
  arrange(start_date, EP_game_id)
```

#############################################################################
#############################################################################

## 3 CREATING A ROLLING WINDOW

Utilising a rolling window; takes a subset of the data and moves the window across the dataset with the window size constant(ish). The window slides forward by an observation at each step.

In this, an 'observation' is a tournament, so the window won't be necessarily have the same number of datapoints but will have the same number of tournaments.

This allows for the dynamic updating of the model as new data becomes available.

```{r}
unique_tournaments <- unique(pairwise_matches$EP_game_id)
length(unique_tournaments)
```

This function seems to be working.

```{r}
create_rolling_windows <- function(data, tournaments, window_size) {
  n <- length(tournaments)
  windows <- list()
  
  for (start in 1:(n - window_size + 1)) {
    end <- start + window_size - 1
    window_tournaments <- tournaments[start:end]
    window_data <- data[data$EP_game_id %in% window_tournaments, ]
    windows[[length(windows) + 1]] <- window_data
  }
  
  return(windows)
}

# window size (number of tournaments per window)
window_size <- 80

windows <- create_rolling_windows(pairwise_matches, unique_tournaments, window_size)

## make sure function is working as expected

length(windows)
length(unique(windows[[4]]$EP_game_id))

```

Create a window dataframe.

```{r}
create_window_summary <- function(windows) {
  summary_df <- data.frame(window_number = integer(),
                           final_EP_game_id = integer(),
                           final_tournament = character(),
                           start_data = character(),
                           major = integer(),
                           stringsAsFactors = FALSE)
  
  for (i in 1:(length(windows) - 1)) {
    next_window <- windows[[i + 1]]
    final_EP_game_id <- tail(next_window$EP_game_id, n = 1)  
    final_tournament <- tail(next_window$tournament, n = 1)  
    final_start_date <- tail(next_window$start_date, n = 1)  
    final_major <- tail(next_window$major, n = 1)  
    
    summary_df <- rbind(summary_df, data.frame(window_number = i,
                                               next_EP_game_id = final_EP_game_id,
                                               next_tournament = final_tournament,
                                               start_date = final_start_date,
                                               major = final_major))
  }
  
  return(summary_df)
}

window_summary <- create_window_summary(windows)
```

## 4 PREPARING DATA FOR MODEL

## 5 TRAINING MODEL

First, define the time-decay factor. how to choose lambda?

```{r}
decay_function <- function(time_diff, lambda) {
    return(exp(-lambda * time_diff))
}
```



In order to implement a time-decay factor, need to append a column with the upcoming tournamnet date.

```{r}
add_time_decay_factor <- function(window_data, window_summary, window_number, lambda = 0.25) {
    # check if the window_number exists in window_summary
    if (window_number > nrow(window_summary)) {
        # skip this window as it isn't used for prediction
        return(window_data)
    }
    
    # get the current tournament start date and name from the window_summary
    upcoming_tournament_date <- window_summary$start_date[window_summary$window_number == window_number]
    upcoming_tournament_name <- window_summary$next_tournament[window_summary$window_number == window_number]
    
    # calculate time differences between the match date and the upcoming tournament start date
    window_data$time_diff <- as.numeric(difftime(upcoming_tournament_date, window_data$start_date, units = "days"))
    
    # apply the decay function
    window_data$time_decay_factor <- decay_function(window_data$time_diff, lambda)
    
    window_data$upcoming_tournament <- upcoming_tournament_date
    window_data$upcoming_tournament_name <- upcoming_tournament_name
    
    return(window_data)
}

# apply the function to each window, skipping those without predictions
windows_decay <- lapply(seq_along(windows), function(i) {
    if (i > nrow(window_summary)) {
        return(windows[[i]])
    } else {
        return(add_time_decay_factor(windows[[i]], window_summary, i))
    }
})

```

Creating adjacency matrices, make sure to skip final window as not predicitve and avoids errors.

```{r}
get_adjacency_matrix_with_decay <- function(window_data) {
    players <- unique(c(window_data$player1, window_data$player2))
    n <- length(players)
    player_indices <- setNames(1:n, players)
  
    adj_matrix <- matrix(0, nrow = n, ncol = n)
    rownames(adj_matrix) <- players
    colnames(adj_matrix) <- players

    for (i in 1:nrow(window_data)) {
        # retrieve the time-decay factor
        decay_factor <- window_data$time_decay_factor[i]
        
        # check if decay_factor is missing or NA
        if (is.na(decay_factor) || length(decay_factor) == 0) {
            warning(paste("Skipping row", i, ": decay_factor is missing or NA"))
            next  # Skip this iteration if decay_factor is missing
        }
        
        # determine the winner and loser
        if (window_data$score1[i] == 1) {  # player1 won
            winner <- window_data$player1[i]
            loser <- window_data$player2[i]
        } else {  # player2 won
            winner <- window_data$player2[i]
            loser <- window_data$player1[i]
        }

        # update the adjacency matrix using the decay factor
        adj_matrix[loser, winner] <- adj_matrix[loser, winner] + decay_factor
    }

    return(adj_matrix)
}


```

Generating for each window, skipping last:

```{r}
# generate adjacency matrices for each window, skipping the last window
adj_matrices <- lapply(seq_along(windows_decay), function(i) {
    if (i > 68) {
        warning(paste("Skipping window", i, ": No prediction available"))
        return(NULL)  # Skip the final window or any window without a prediction
    }
    return(get_adjacency_matrix_with_decay(windows_decay[[i]]))
})

```

Computing PageRanks:
```{r}
compute_pagerank <- function(adj_matrix) {
    g <- graph_from_adjacency_matrix(adj_matrix, mode = "directed", weighted = TRUE)
    pr <- page_rank(g, damping = 0.85)$vector
    return(pr)
}
```


Getting Results:

```{r}
pagerank_results <- lapply(adj_matrices, function(adj_matrix) {
    compute_pagerank(adj_matrix)
})
print(pagerank_results[[2]])
```

Creating a datframe for player ranks:

```{r}
# get the unique list of players across all windows
all_players <- unique(unlist(lapply(windows, function(w) unique(c(w$player1, w$player2)))))

pagerank_df <- data.frame(matrix(NA, nrow = length(all_players), ncol = length(windows_decay)))
rownames(pagerank_df) <- all_players
colnames(pagerank_df) <- paste0("Window_", seq_along(windows_decay))

for (i in seq_along(pagerank_results)) {

  pr_scores <- pagerank_results[[i]]
  
  # match the players in the PageRank results to the rows in the dataframe
  pagerank_df[match(names(pr_scores), rownames(pagerank_df)), i] <- pr_scores
}

```

## 7. PREDICTION ANALYSIS

Creating a predictive dataframe:

```{r}
predictive_dataframe_pr2 <- merge(pairwise_matches, window_summary, 
                       by.x = "EP_game_id", by.y = "next_EP_game_id", all.x = TRUE) %>%
  select(window_number, start_date.x, EP_game_id, tournament, player1, player2, final_position_A, final_position_B, score1, score2, times_played_A, times_played_B, is.home_A, is.home_B, major.x) %>%
  na.omit() %>% # gets rid of all data prior to window 1
  rename(major = major.x, start_date = start_date.x, player1_win = score1)

```

Formatting dataframe into long format so can add strengths:

```{r}
# assuming player names are rownames
pagerank_df <- pagerank_df %>% 
  rownames_to_column(var = "Player")

# reshape the data
pagerank_long <- pagerank_df %>%
  pivot_longer(cols = starts_with("Window_"),  # Use starts_with to gather all Window columns
               names_to = "Window", 
               values_to = "Rank")

# extract the window number from the 'Window' column
pagerank_long$Window <- as.numeric(sub("Window_", "", pagerank_long$Window))
```

Adding strengths:

```{r}
predictive_dataframe_pr2 <- predictive_dataframe_pr2 %>%
  left_join(pagerank_long %>% select(Player, Window, Rank), 
            by = c("player1" = "Player", "window_number" = "Window")) %>%
  rename(ranking_player1 = Rank) %>%
  left_join(pagerank_long %>% select(Player, Window, Rank), 
            by = c("player2" = "Player", "window_number" = "Window")) %>%
  rename(ranking_player2 = Rank)

```

Adding a column for predicted winners: P(A beats B) = rank_A / (rank_A + rank_B)

```{r}
predictive_dataframe_pr2 <- predictive_dataframe_pr2 %>%
  mutate(prob_player1_win = ranking_player1/(ranking_player1 + ranking_player2),
         prob_player2_win = ranking_player2/(ranking_player1 + ranking_player2))
```

```{r}
predictive_dataframe_pr2 <- predictive_dataframe_pr2 %>%
  select(-start_date, -final_position_A, -final_position_B, -times_played_A, -times_played_B, -is.home_A, -is.home_B, -score2)

```

Calculating Accuracy:

```{r}
# get rid of all non-predictive data
predictive_dataframe_pr2 <- predictive_dataframe_pr2 %>%
  drop_na()

correct_predictions <- predictive_dataframe_pr2 %>%
  mutate(predicted_win = ifelse(prob_player1_win > 0.5, 1, 0)) %>%
  summarize(accuracy = mean(predicted_win == player1_win))

print(correct_predictions)
```

```{r}
dim(predictive_dataframe_pr2)
```

Calculate log loss:

```{r}
log_loss <- predictive_dataframe_pr2 %>%
  mutate(log_loss = - (player1_win * log(prob_player1_win) + (1 - player1_win) * log(1 - prob_player1_win))) %>%
  summarize(mean_log_loss = mean(log_loss))

print(log_loss)
```
Calculating Brier Score:

```{r}
brier_score <- predictive_dataframe_pr2 %>%
  mutate(brier_score = (prob_player1_win - player1_win)^2) %>%
  summarize(mean_brier_score = mean(brier_score))

print(brier_score)
```
Accuracy per window:

```{r}
predictive_dataframe_pr2 <- predictive_dataframe_pr2 %>%
    mutate(predicted_win = ifelse(prob_player1_win > 0.5, 1, 0))

accuracy_per_window_pr2 <- predictive_dataframe_pr2 %>%
    group_by(window_number) %>%  # Group by the window number
    summarize(accuracy = mean(predicted_win == player1_win))

print(accuracy_per_window_pr2)

ggplot(accuracy_per_window_pr2, aes(x = window_number, y = accuracy)) +
    geom_line(color = "blue") +
    geom_point(color = "red") +
    labs(
        title = "Accuracy per Window",
        x = "Window Number",
        y = "Accuracy"
    ) +
    theme_minimal()
```

## PREDICTING AGAINST THE BOOKS

## 8. PREDICTING AGAINST THE BOOKS

```{r}
# merge the odds_data with the original predictive_dataframe
merged_original <- predictive_dataframe_pr2 %>%
  left_join(odds_data, by = c("EP_game_id", "player1", "player2")) %>%
  rename(
    golfer1_win_orig = golfer1_win,
    golfer2_win_orig = golfer2_win
  )

# swap player1 and player2 in odds_data for the merge
odds_data_swapped <- odds_data %>%
  rename(
    player1 = player2,
    player2 = player1,
    golfer1_win = golfer2_win,
    golfer2_win = golfer1_win
  )

# merge the odds_data with the predictive_dataframe with swapped order
merged_swapped <- predictive_dataframe_pr2 %>%
  left_join(odds_data_swapped, by = c("EP_game_id", "player1", "player2")) %>%
  rename(
    golfer1_win_swapped = golfer1_win,
    golfer2_win_swapped = golfer2_win
  )

# combine the results from the original and swapped merges
predictive_odds_data_pr2 <- merged_original %>%
  left_join(merged_swapped, by = c("EP_game_id", "player1", "player2", "player1_win",   "ranking_player1", "ranking_player2", "major"))

# coalesce the columns to handle NA values and get the correct golfer1_win and golfer2_win
predictive_odds_data_pr2 <- predictive_odds_data_pr2 %>%
  mutate(
    tournament = coalesce(tournament.x.x, tournament.x.y, tournament.y.x, tournament.y.y),
    golfer1_win = coalesce(golfer1_win_orig, golfer1_win_swapped),
    golfer2_win = coalesce(golfer2_win_orig, golfer2_win_swapped),
    prob_player1_win = coalesce(prob_player1_win.x, prob_player1_win.y),
    prob_player2_win = coalesce(prob_player2_win.x, prob_player2_win.y)
  ) %>%
  select(-golfer1_win_orig, -golfer2_win_orig, -golfer1_win_swapped, -golfer2_win_swapped, -tournament.x.x, -tournament.x.y, -tournament.y.x, -tournament.y.y, -prob_player1_win.x, -prob_player1_win.y, -prob_player2_win.x, -prob_player2_win.y, window_number.x) %>%
  rename(window_number = window_number.x)
```

Selecting relevant columns and deleting data used in the first window:

```{r}
predictive_odds_data_pr2 <- predictive_odds_data_pr2 %>%
  select(EP_game_id, tournament, player1, player2, golfer1_win, golfer2_win, prob_player1_win, prob_player2_win, player1_win, window_number, major)

predictive_odds_data_pr2 <- predictive_odds_data_pr2 %>%
  drop_na()
```


Calculating accuracy of these predictions:

```{r}
correct_predictions <- predictive_odds_data_pr2 %>%
  mutate(predicted_win = ifelse(prob_player1_win > 0.5, 1, 0)) %>%
  summarize(accuracy = mean(predicted_win == player1_win))

print(correct_predictions)
```
Calculate log loss:

```{r}
log_loss <- predictive_odds_data_pr2 %>%
  mutate(log_loss = - (player1_win * log(prob_player1_win) + (1 - player1_win) * log(1 - prob_player1_win))) %>%
  summarize(mean_log_loss = mean(log_loss))

print(log_loss)
```
Calculating Brier Score:

```{r}
brier_score <- predictive_odds_data_pr2 %>%
  mutate(brier_score = (prob_player1_win - player1_win)^2) %>%
  summarize(mean_brier_score = mean(brier_score))

print(brier_score)
```

